---
title: "qual_analysis"
author: "Matthew Davis"
date: "2024-11-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup
## Libraries
```{r, message=FALSE}
library(tidyverse)
library(data.table)
library(pbapply)
library(viridis)
library(ggrepel)
library(scales)
```

## Functions
```{r}
mut_per_qual <- function(mut_table, qual_range = 20:40){
  
  muts <- pblapply(qual_range, function(i) {
    qual_muts <- mut_table %>%
      filter(QUAL >= i) %>%
      group_by(SOURCE) %>%
      summarise(qvalue = i, qual_muts = n(), .groups = 'drop')
    
    med_muts <- mut_table %>%
      filter(medQUAL >= i) %>%
      group_by(SOURCE) %>%
      summarise(qvalue = i, medqual_muts = n(), .groups = 'drop')
    
    inner_join(qual_muts, med_muts, by = c("SOURCE", "qvalue"))
  })
  
  # Combine all results into a single data frame if desired
  mut_table <- bind_rows(muts)
  
  return(mut_table)
}
```


## Read in mutation data
```{r}
mutations <- fread("./Data/VCFs/Parsed/mutations_strict.tsv")
```

# Prep data 
## Filter for SNPs
```{r}
snps <- mutations %>%
  filter(SNP == TRUE, MAPP == TRUE)
indels <- mutations %>%
  filter(SNP == FALSE, MAPP == TRUE)
```

## SNP number per quality score
```{r}
qual_table <- mut_per_qual(snps)
qual_table_indels <- mut_per_qual(indels)
```

## Prep table for plotting
### Organize levels and add defining columns
```{r}
levels <- c( "tree2", "ref_chandler", "cr2_hifi",
            "cr2_16A2", "cr2_16A1", "cr2_17A2",
            "cr2_17A1", "cr2_18A1", "cr2_16A",
            "cr2_15A1", "cr2_15A", "cr2_18A",
            "cr21_2", "cr21_1", "cr2_12A",
            "cr2_13A", "cr2_11A",
            "cr85", "cr10", "cr13", "cr22")

qual_table$SOURCE <- factor(qual_table$SOURCE, levels = levels)

qual_table <- qual_table %>%
  mutate(TYPE = case_when(
    SOURCE %in% c("ref_chandler", "tree2", "cr2_hifi") ~ "HiFi",
    SOURCE %in% c("cr21_1", "cr21_2", "cr2_11A", "cr2_12A", "cr2_13A", "cr2_15A", "cr2_15A1", "cr2_16A", "cr2_16A1",
                  "cr2_16A2", "cr2_17A1", "cr2_17A2", "cr2_18A", "cr2_18A1") ~ "Short Read Embryo",
    SOURCE %in% c("cr85", "cr10", "cr13", "cr22") ~ "Short Read Shoot")) %>%
  mutate(TYPE = factor(TYPE, levels = c("HiFi", "Short Read Embryo", "Short Read Shoot")))

qual_table <- qual_table %>%
  mutate(CULTURETYPE = case_when(
    SOURCE %in% c("ref_chandler", "tree2") ~ "tree",
    SOURCE %in% c("cr2_hifi", "cr21_1", "cr21_2", "cr2_11A", "cr2_12A", "cr2_13A", "cr2_15A",
                  "cr2_15A1", "cr2_16A", "cr2_16A1", "cr2_16A2", "cr2_17A1", "cr2_17A2",
                  "cr2_18A", "cr2_18A1") ~ "embryo",
    SOURCE %in% c("cr85", "cr10", "cr13", "cr22") ~ "shoot")) %>%
  mutate(SEQTYPE = case_when(
    SOURCE %in% c("ref_chandler", "tree2", "cr2_hifi") ~ "long_read",
    SOURCE %in% c("cr21_1", "cr21_2", "cr2_11A", "cr2_12A", "cr2_13A", "cr2_15A", "cr2_15A1", 
                  "cr2_16A", "cr2_16A1","cr2_16A2", "cr2_17A1", "cr2_17A2", "cr2_18A",
                  "cr2_18A1", "cr85", "cr10", "cr13", "cr22") ~ "short_read"))
```

### Add plotting specific annotations
```{r}
# Add labels to annotate
qual_table <- qual_table %>%
  mutate(
    label = ifelse(SOURCE == "cr2_hifi" & qvalue %in% c(20, 40), as.character(paste("Quality\n", qvalue)), NA)  # Create labels for qvalues 20 and 40
  )

# Rescale qvalues for plotting and add gradient
qual_table <- qual_table %>%
  mutate(
    qvalue_rescaled = rescale(qvalue, to = c(0, 1)),  # Rescale qvalue to the range [0, 1]
    color_gradient = case_when(
      CULTURETYPE == "tree" ~ gradient_n_pal(c("grey80", "#B4C5B4", "darkseagreen4"))(qvalue_rescaled),
      CULTURETYPE == "shoot" ~ gradient_n_pal(c("grey80", "#D9F6D9","darkseagreen2"))(qvalue_rescaled),
      CULTURETYPE == "embryo" ~ gradient_n_pal(c("grey80", "#FFE092","goldenrod1"))(qvalue_rescaled)
    )
  )
```

# Plot data
## Scale color virdis
```{r}
qual_table %>%
  ggplot(aes(x = SOURCE, y = medqual_muts, col = qvalue)) +
  geom_point() +
  scale_color_viridis(name = "Quality\nThreshold", option = "cividis") +
  labs(y = "Number of Mutations", x = "Individual") +
  facet_grid(~TYPE, scales = "free", space = "free") +
  scale_x_discrete(labels = c("ref_chandler" = "Reference", "tree2" = "Tree",
                              "cr2_hifi" = "E HiFi", "cr21_1" = "E 11", "cr21_2" = "E 10",
                              "cr2_11A" = "E 14", "cr2_12A" = "E 12", "cr2_13A" = "E 13",
                              "cr2_15A" = "E 8", "cr2_15A1" = "E 7", "cr2_16A" = "E 6",
                              "cr2_16A1" = "E 2", "cr2_16A2" = "E 1", "cr2_17A1" = "E 4",
                              "cr2_17A2" = "E 3", "cr2_18A" = "E 9", "cr2_18A1" = "E 5",
                              "cr85" = "S 1985", "cr10" = "S 2010",
                              "cr13" = "S 2013", "cr22" = "S 2022")) +
  theme_classic(base_size = 6) +
  theme(legend.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
        axis.title.x = element_blank(),
        axis.title.y = element_text(face = "bold"),
        strip.background = element_blank(),
        strip.text = element_text(face = "bold"),
        legend.position = "bottom",
        legend.key.size = unit(0.2, "cm"))
# ggsave("./Figures/Quality_Analysis/medquality_comparison_snps.pdf", height = 2, width = 4)
# ggsave("./Figures/Quality_Analysis/medquality_comparison_snps.png", height = 2, width = 4)
```

## Gradients based on culture type
```{r}
qual_table %>%
  ggplot(aes(x = SOURCE, y = medqual_muts)) +
  geom_point(aes(color = color_gradient)) +
  scale_color_identity(guide = "legend") +
  labs(y = "Number of Mutations", x = "Individual", color = "q-value") +
  facet_grid(~TYPE, scales = "free", space = "free") +
  scale_x_discrete(
    labels = c(
      "ref_chandler" = "Reference", "tree2" = "Tree",
      "cr2_hifi" = "E HiFi", "cr21_1" = "E 11", "cr21_2" = "E 10",
      "cr2_11A" = "E 14", "cr2_12A" = "E 12", "cr2_13A" = "E 13",
      "cr2_15A" = "E 8", "cr2_15A1" = "E 7", "cr2_16A" = "E 6",
      "cr2_16A1" = "E 2", "cr2_16A2" = "E 1", "cr2_17A1" = "E 4",
      "cr2_17A2" = "E 3", "cr2_18A" = "E 9", "cr2_18A1" = "E 5",
      "cr85" = "S 1985", "cr10" = "S 2010",
      "cr13" = "S 2013", "cr22" = "S 2022"
    ),
    #expand = c(.1,0)
  ) +
  geom_text_repel(
    data = filter(qual_table, label == "Quality\n 20"),  # Replace with the condition for the first label
    aes(label = label),
    na.rm = TRUE,
    size = 1,
    nudge_y = 3 ,
    nudge_x = -.6# Adjust nudge direction for Label1
  ) +
  geom_text_repel(
    data = filter(qual_table, label == "Quality\n 40"),  # Replace with the condition for the second label
    aes(label = label),
    na.rm = TRUE,
    size = 1,
    nudge_y = -.3,
    nudge_x = -.1# Adjust nudge direction for Label2
  ) +
  theme_classic(base_size = 6) +
  theme(
    legend.title = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 35, vjust = 1, hjust = 1),
    axis.title.x = element_blank(),
    strip.background = element_blank(),
    #strip.text = element_text(face = "bold"),
    legend.position = "none",
    #panel.spacing = unit(.5, "cm"),
    plot.margin = unit(c(0, 0, 0, 0), "cm")
  )
ggsave("./Figures/Quality_Analysis/medquality_comparison_snps.pdf", height = 1.5, width = 4)
ggsave("./Figures/Quality_Analysis/medquality_comparison_snps.png", height = 1.5, width = 4)
```

## Flipped axes
### Create new levels and geom_text labels
```{r}
# Levels
levels <- c( "tree2", "ref_chandler", "cr22", "cr13", "cr10", "cr85",
            "cr2_11A", "cr2_13A", "cr2_12A",  "cr21_1", "cr21_2",
            "cr2_18A", "cr2_15A", "cr2_15A1", "cr2_hifi", "cr2_16A",
            "cr2_18A1", "cr2_17A1", "cr2_17A2", "cr2_16A1", "cr2_16A2")

qual_table$SOURCE <- factor(qual_table$SOURCE, levels = levels)

# Text
# Add labels to annotate
qual_table <- qual_table %>%
  mutate(
    label = ifelse(SOURCE == "cr2_16A2" & qvalue %in% c(20, 40), as.character(paste("Quality\n", qvalue)), NA)  # Create labels for qvalues 20 and 40
  )

# Rescale qvalues for plotting and add gradient
qual_table <- qual_table %>%
  mutate(
    qvalue_rescaled = rescale(qvalue, to = c(0, 1)),  # Rescale qvalue to the range [0, 1]
    color_gradient = case_when(
      CULTURETYPE == "tree" ~ gradient_n_pal(c("grey80", "#B4C5B4", "darkseagreen4"))(qvalue_rescaled),
      CULTURETYPE == "shoot" ~ gradient_n_pal(c("grey80", "#D9F6D9","darkseagreen2"))(qvalue_rescaled),
      CULTURETYPE == "embryo" ~ gradient_n_pal(c("grey80", "#FFE092","goldenrod1"))(qvalue_rescaled)
    )
  )
```

## Plot flipped axes for SNPs
```{r}
qual_table %>%
  ggplot(aes(x = medqual_muts, y = SOURCE)) +
  geom_point(aes(color = color_gradient), size = 0.8) +
  scale_color_identity(guide = "legend") +
  labs(x = "Number of SBS", color = "q-value") +
  #facet_grid(~TYPE, scales = "free", space = "free") +
  scale_y_discrete(
    labels = c(
      "ref_chandler" = "Reference", "tree2" = "Tree",
      "cr2_hifi" = "E HiFi", "cr21_1" = "E 11", "cr21_2" = "E 10",
      "cr2_11A" = "E 14", "cr2_12A" = "E 12", "cr2_13A" = "E 13",
      "cr2_15A" = "E 8", "cr2_15A1" = "E 7", "cr2_16A" = "E 6",
      "cr2_16A1" = "E 2", "cr2_16A2" = "E 1", "cr2_17A1" = "E 4",
      "cr2_17A2" = "E 3", "cr2_18A" = "E 9", "cr2_18A1" = "E 5",
      "cr85" = "S 1985", "cr10" = "S 2010",
      "cr13" = "S 2013", "cr22" = "S 2022"
    )
  ) +
  scale_x_continuous(limits = c(0,3200), breaks = c(0, 1000, 2000, 3000)) +
  geom_text_repel(
    data = filter(qual_table, label == "Quality\n 20"),  # Replace with the condition for the first label
    aes(label = label),
    na.rm = TRUE,
    size = 1,
    segment.color = NA,
    y = 22,
    x = 3200
  ) +
  geom_text_repel(
    data = filter(qual_table, label == "Quality\n 40"),  # Replace with the condition for the second label
    aes(label = label),
    na.rm = TRUE,
    size = 1,
    segment.color = NA,
    y = 22,
    x = 1300
  ) +
  theme_classic(base_size = 6) +
  theme(
    legend.title = element_text(hjust = 0.5),
    axis.text.y = element_blank(),
    axis.title.y = element_blank(),
    axis.line.y = element_blank(),
    axis.ticks.y = element_blank(),
    strip.background = element_blank(),
    legend.position = "none",
    panel.background = element_rect(fill = NA, color = NA),
    plot.background = element_rect(fill = NA, color = NA),
    plot.margin = unit(c(0, 0.1, 0, 0), "cm")
  )
ggsave("./Figures/Quality_Analysis/medquality_comparison_snps_long.pdf", height = 2.55, width = .9)
ggsave("./Figures/Quality_Analysis/medquality_comparison_snps_long.png", height = 2.55, width = .9)
```

# Plotting INDELS 
## Organize levels and add defining columns
```{r}
# Levels
levels <- c( "tree2", "ref_chandler", "cr22", "cr13", "cr10", "cr85",
            "cr2_11A", "cr2_13A", "cr2_12A",  "cr21_1", "cr21_2",
            "cr2_18A", "cr2_15A", "cr2_15A1", "cr2_hifi", "cr2_16A",
            "cr2_18A1", "cr2_17A1", "cr2_17A2", "cr2_16A1", "cr2_16A2")

qual_table_indels$SOURCE <- factor(qual_table_indels$SOURCE, levels = levels)

qual_table_indels <- qual_table_indels %>%
  mutate(TYPE = case_when(
    SOURCE %in% c("ref_chandler", "tree2", "cr2_hifi") ~ "HiFi",
    SOURCE %in% c("cr21_1", "cr21_2", "cr2_11A", "cr2_12A", "cr2_13A", "cr2_15A", "cr2_15A1", "cr2_16A", "cr2_16A1",
                  "cr2_16A2", "cr2_17A1", "cr2_17A2", "cr2_18A", "cr2_18A1") ~ "Short Read Embryo",
    SOURCE %in% c("cr85", "cr10", "cr13", "cr22") ~ "Short Read Shoot")) %>%
  mutate(TYPE = factor(TYPE, levels = c("HiFi", "Short Read Embryo", "Short Read Shoot")))

qual_table_indels <- qual_table_indels %>%
  mutate(CULTURETYPE = case_when(
    SOURCE %in% c("ref_chandler", "tree2") ~ "tree",
    SOURCE %in% c("cr2_hifi", "cr21_1", "cr21_2", "cr2_11A", "cr2_12A", "cr2_13A", "cr2_15A",
                  "cr2_15A1", "cr2_16A", "cr2_16A1", "cr2_16A2", "cr2_17A1", "cr2_17A2",
                  "cr2_18A", "cr2_18A1") ~ "embryo",
    SOURCE %in% c("cr85", "cr10", "cr13", "cr22") ~ "shoot")) %>%
  mutate(SEQTYPE = case_when(
    SOURCE %in% c("ref_chandler", "tree2", "cr2_hifi") ~ "long_read",
    SOURCE %in% c("cr21_1", "cr21_2", "cr2_11A", "cr2_12A", "cr2_13A", "cr2_15A", "cr2_15A1", 
                  "cr2_16A", "cr2_16A1","cr2_16A2", "cr2_17A1", "cr2_17A2", "cr2_18A",
                  "cr2_18A1", "cr85", "cr10", "cr13", "cr22") ~ "short_read"))
```

## Add plotting specific annotations
```{r}
# Text
# Add labels to annotate
qual_table_indels <- qual_table_indels %>%
  mutate(
    label = ifelse(SOURCE == "cr2_16A2" & qvalue %in% c(20, 40), as.character(paste("Quality\n", qvalue)), NA)  # Create labels for qvalues 20 and 40
  )

# Rescale qvalues for plotting and add gradient
qual_table_indels <- qual_table_indels %>%
  mutate(
    qvalue_rescaled = rescale(qvalue, to = c(0, 1)),  # Rescale qvalue to the range [0, 1]
    color_gradient = case_when(
      CULTURETYPE == "tree" ~ gradient_n_pal(c("grey80", "#B4C5B4", "darkseagreen4"))(qvalue_rescaled),
      CULTURETYPE == "shoot" ~ gradient_n_pal(c("grey80", "#D9F6D9","darkseagreen2"))(qvalue_rescaled),
      CULTURETYPE == "embryo" ~ gradient_n_pal(c("grey80", "#FFE092","goldenrod1"))(qvalue_rescaled)
    )
  )
```

## Plot indels
```{r}
qual_table_indels %>%
  ggplot(aes(x = medqual_muts, y = SOURCE)) +
  geom_point(aes(color = color_gradient), size = 0.8) +
  scale_color_identity(guide = "legend") +
  labs(x = "Number of INDELs", color = "q-value") +
  #facet_grid(~TYPE, scales = "free", space = "free") +
  scale_y_discrete(
    labels = c(
      "ref_chandler" = "Reference", "tree2" = "Tree",
      "cr2_hifi" = "E HiFi", "cr21_1" = "E 11", "cr21_2" = "E 10",
      "cr2_11A" = "E 14", "cr2_12A" = "E 12", "cr2_13A" = "E 13",
      "cr2_15A" = "E 8", "cr2_15A1" = "E 7", "cr2_16A" = "E 6",
      "cr2_16A1" = "E 2", "cr2_16A2" = "E 1", "cr2_17A1" = "E 4",
      "cr2_17A2" = "E 3", "cr2_18A" = "E 9", "cr2_18A1" = "E 5",
      "cr85" = "S 1985", "cr10" = "S 2010",
      "cr13" = "S 2013", "cr22" = "S 2022"
    )
  ) +
  scale_x_continuous(limits = c(0,3200), breaks = c(0, 1000, 2000, 3000)) +
  geom_text_repel(
    data = filter(qual_table_indels, label == "Quality\n 20"),  # Replace with the condition for the first label
    aes(label = label),
    na.rm = TRUE,
    size = 1,
    segment.color = NA,
    y = 22,
    x = 2800
  ) +
  geom_text_repel(
    data = filter(qual_table_indels, label == "Quality\n 40"),  # Replace with the condition for the second label
    aes(label = label),
    na.rm = TRUE,
    size = 1,
    segment.color = NA,
    y = 22,
    x = 500
  ) +
  theme_classic(base_size = 6) +
  theme(
    legend.title = element_text(hjust = 1),
    axis.text.y = element_blank(),
    axis.title.y = element_blank(),
    axis.line.y = element_blank(),
    axis.ticks.y = element_blank(),
    strip.background = element_blank(),
    legend.position = "none",
    panel.background = element_rect(fill = NA, color = NA),
    plot.background = element_rect(fill = NA, color = NA),
    plot.margin = unit(c(0, 0.1, 0, 0), "cm")
  )
ggsave("./Figures/Quality_Analysis/medquality_comparison_indels_long.pdf", height = 2.55, width = .9)
ggsave("./Figures/Quality_Analysis/medquality_comparison_indels_long.png", height = 2.55, width = .9)
```

