---
title: "te_near_genes"
author: "Matthew Davis"
date: "2025-01-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Setup
## Libraries
```{r, message=FALSE}
library(ggenomics)
library(parallel)
library(foreach)
library(doParallel)
library(ggpubr)
```

## Functions
```{r}
in_near_gene <- function(sv, genes){
  
  for (i in 1:nrow(sv)) {
    sv_row <- sv[i]
    
    relevant_genes <- genes[CHROM == sv_row$CHROM & HAP == sv_row$HAP]
    
    abs_diff_start <- abs(relevant_genes$START - sv_row$POS)
    abs_diff_stop <- abs(relevant_genes$STOP - sv_row$POS)
    
    is_genic <- any(sv_row$POS >= relevant_genes$START & sv_row$POS <= relevant_genes$STOP)
    
    sv[i, `:=`(CLOSEST_START = min(abs_diff_start), CLOSEST_STOP = min(abs_diff_stop), GENIC = is_genic)]
  }
}

in_features <- function(sv, gff){
  
  for (i in 1:nrow(sv)) {
    sv_row <- sv[i]
    
    relevant_5utr <- gff[CHROM == sv_row$CHROM & HAP == sv_row$HAP & TYPE == "5UTR"]
    relevant_3utr <- gff[CHROM == sv_row$CHROM & HAP == sv_row$HAP & TYPE == "3UTR"]
    relevant_exon <- gff[CHROM == sv_row$CHROM & HAP == sv_row$HAP & TYPE == "exon"]
    relevant_intron <- gff[CHROM == sv_row$CHROM & HAP == sv_row$HAP & TYPE == "intron"]
    relevant_inter <- gff[CHROM == sv_row$CHROM & HAP == sv_row$HAP & TYPE == "intergenic"]
    relevant_up <- gff[CHROM == sv_row$CHROM & HAP == sv_row$HAP & TYPE == "upstream"]
    relevant_down <- gff[CHROM == sv_row$CHROM & HAP == sv_row$HAP & TYPE == "downstream"]
    
    is_5utr <- any(sv_row$POS >= relevant_5utr$START & sv_row$POS <= relevant_5utr$STOP)
    is_3utr <- any(sv_row$POS >= relevant_3utr$START & sv_row$POS <= relevant_3utr$STOP)
    is_exon <- any(sv_row$POS >= relevant_exon$START & sv_row$POS <= relevant_exon$STOP)
    is_intron <- any(sv_row$POS >= relevant_intron$START & sv_row$POS <= relevant_intron$STOP)
    is_inter <- any(sv_row$POS >= relevant_inter$START & sv_row$POS <= relevant_inter$STOP)
    is_up <- any(sv_row$POS >= relevant_up$START & sv_row$POS <= relevant_up$STOP)
    is_down <- any(sv_row$POS >= relevant_down$START & sv_row$POS <= relevant_down$STOP)
    
    
    sv[i, `:=`(UTR5 = is_5utr, UTR3 = is_3utr, EXON = is_exon, INTRON = is_intron, INTERGENIC = is_inter, UPSTREAM = is_up, DOWNSTREAM = is_down)]
  }
}

sim_in_near_gene <- function(sv, genes, cores = 8){
  
  # Set up parallel
  num_cores <- cores
  cl <- makeCluster(num_cores)
  registerDoParallel(cl)
  
  # Create a list to store the outputs
  random_sv_list <- foreach(iteration = 1:10000, .packages = 'data.table') %dopar% {
    
    # Randomize the simulation position for each iteration
    sv[, SIM_POS := sample(1:CHROM_LENGTH, 1), by = .I]
    
    # Create columns
    sv[, `:=`(SIM_START = NA_integer_, SIM_STOP = NA_integer_, SIM_GENIC = FALSE)]
    
    # Calculate the closest gene positions
    sv[, `:=`(
      SIM_START = sapply(1:nrow(sv), function(i) {
        sv_row <- sv[i]
        relevant_genes <- genes[CHROM == sv_row$CHROM & HAP == sv_row$HAP]
        abs_diff_start <- abs(relevant_genes$START - sv_row$SIM_POS)
        min(abs_diff_start)
      }),
      SIM_STOP = sapply(1:nrow(sv), function(i) {
        sv_row <- sv[i]
        relevant_genes <- genes[CHROM == sv_row$CHROM & HAP == sv_row$HAP]
        abs_diff_stop <- abs(relevant_genes$STOP - sv_row$SIM_POS)
        min(abs_diff_stop)
      })
    )]
    
    # Check if SIM_POS falls within any START-STOP range
    sv[, SIM_GENIC := sapply(1:nrow(sv), function(i) {
      sv_row <- sv[i]
      relevant_genes <- genes[CHROM == sv_row$CHROM & HAP == sv_row$HAP]
      any(sv_row$SIM_POS >= relevant_genes$START & sv_row$SIM_POS <= relevant_genes$STOP)
    })]
    
    # Return the result of each iteration
    return(copy(sv))
  }
  
  # Stop the cluster after the computation is done
  stopCluster(cl)
  
  return(random_sv_list)
}

sim_list_means <- function(sim_list){
  
  # Create empty variable for loop
  sim_means <- c()
  
  for (i in 1:length(sim_list)) {
    
    sim_means[[i]] <- sim_list[[i]] %>%
      group_by(SOURCE, TE) %>%
      summarise(
        NUMBER = n(),
        MEAN_SIMSTART = mean(SIM_START),
        MEAN_SIMSTOP = mean(SIM_STOP),
        TOTAL_SIM_GENIC = sum(SIM_GENIC),
        PROP_SIM_GENIC = TOTAL_SIM_GENIC/NUMBER
      ) %>%
      ungroup()
  }
  
  sim_means_table <- rbindlist(sim_means)
  
  return(sim_means_table)
}

sim_prox_table <- function(sim_means_table){
  
  prox_sim_means <- sim_means_table %>%
    pivot_longer(cols = c(MEAN_SIMSTART, MEAN_SIMSTOP), names_to = "LOCATION", values_to = "DISTANCE")
  setDT(prox_sim_means)
  prox_sim_means[, LOCATION := fifelse(LOCATION == "MEAN_SIMSTART", "start", "stop")]
  prox_sim_means[, UNIQUE := paste(SOURCE, TE, LOCATION, sep = "_")]
  
  return(prox_sim_means)
}

real_prox_table <- function(real_means_table){
  
  prox_real_means <- real_means_table %>%
    pivot_longer(cols = c(MEAN_CLOSESTART, MEAN_CLOSESTOP), names_to = "LOCATION", values_to = "DISTANCE")
  setDT(prox_real_means)
  prox_real_means[, LOCATION := fifelse(LOCATION == "MEAN_CLOSESTART", "start", "stop")]
  prox_real_means[, UNIQUE := paste(SOURCE, TE, LOCATION, sep = "_")]
  
  return(prox_real_means)
}

sim_in_feature <- function(sv, gff) {
  
  for (i in 1:nrow(sv)) {
    sv_row <- sv[i]
    
    relevant_5utr <- gff[CHROM == sv_row$CHROM & HAP == sv_row$HAP & TYPE == "5UTR"]
    relevant_3utr <- gff[CHROM == sv_row$CHROM & HAP == sv_row$HAP & TYPE == "3UTR"]
    relevant_exon <- gff[CHROM == sv_row$CHROM & HAP == sv_row$HAP & TYPE == "exon"]
    relevant_intron <- gff[CHROM == sv_row$CHROM & HAP == sv_row$HAP & TYPE == "intron"]
    relevant_inter <- gff[CHROM == sv_row$CHROM & HAP == sv_row$HAP & TYPE == "intergenic"]
    relevant_up <- gff[CHROM == sv_row$CHROM & HAP == sv_row$HAP & TYPE == "upstream"]
    relevant_down <- gff[CHROM == sv_row$CHROM & HAP == sv_row$HAP & TYPE == "downstream"]
    
    is_5utr_sim <- any(sv_row$SIM_POS >= relevant_5utr$START & sv_row$SIM_POS <= relevant_5utr$STOP)
    is_3utr_sim <- any(sv_row$SIM_POS >= relevant_3utr$START & sv_row$SIM_POS <= relevant_3utr$STOP)
    is_exon_sim <- any(sv_row$SIM_POS >= relevant_exon$START & sv_row$SIM_POS <= relevant_exon$STOP)
    is_intron_sim <- any(sv_row$SIM_POS >= relevant_intron$START & sv_row$SIM_POS <= relevant_intron$STOP)
    is_inter_sim <- any(sv_row$SIM_POS >= relevant_inter$START & sv_row$SIM_POS <= relevant_inter$STOP)
    is_up_sim <- any(sv_row$SIM_POS >= relevant_up$START & sv_row$SIM_POS <= relevant_up$STOP)
    is_down_sim <- any(sv_row$SIM_POS >= relevant_down$START & sv_row$SIM_POS <= relevant_down$STOP)
    
    
    
    sv[i, `:=`(SIM_UTR5 = is_5utr_sim, SIM_UTR3 = is_3utr_sim, SIM_EXON = is_exon_sim, SIM_INTRON = is_intron_sim, SIM_INTERGENIC = is_inter_sim, SIM_UPSTREAM = is_up_sim, SIM_DOWNSTREAM = is_down_sim)]
  }
  return(sv)
}

feature_means <- function(real_sv){
  
  means_table <- real_sv %>%
    group_by(SOURCE, TE) %>%
    summarise(
      NUMBER = n(),
      MEAN_START = mean(CLOSEST_START),
      MEAN_STOP = mean(CLOSEST_STOP),
      TOTAL_GENIC = sum(GENIC),
      PROP_GENIC = TOTAL_GENIC/NUMBER,
      TOTAL_5UTR = sum(UTR5),
      PROP_5UTR = TOTAL_5UTR/NUMBER,
      TOTAL_3UTR = sum(UTR3),
      PROP_3UTR = TOTAL_3UTR/NUMBER,
      TOTAL_EXON = sum(EXON),
      PROP_EXON = TOTAL_EXON/NUMBER,
      TOTAL_INTRON = sum(INTRON),
      PROP_INTRON = TOTAL_INTRON/NUMBER,
      TOTAL_INTERGENIC = sum(INTERGENIC),
      PROP_INTERGENIC = TOTAL_INTERGENIC/NUMBER,
      TOTAL_UPSTREAM = sum(UPSTREAM),
      PROP_UPSTREAM = TOTAL_UPSTREAM/NUMBER,
      TOTAL_DOWNSTREAM = sum(DOWNSTREAM),
      PROP_DOWNSTREAM = TOTAL_DOWNSTREAM/NUMBER
    ) %>%
    ungroup()
  
  return(means_table)
}


sim_feature_list_means <- function(sim_list){
  
  # Create empty variable for loop
  sim_means <- c()
  
  for (i in 1:length(sim_list)) {
    
    sim_means[[i]] <- sim_list[[i]] %>%
      group_by(SOURCE, TE) %>%
      summarise(
        NUMBER = n(),
        MEAN_SIMSTART = mean(SIM_START),
        MEAN_SIMSTOP = mean(SIM_STOP),
        TOTAL_SIM_GENIC = sum(SIM_GENIC),
        PROP_SIM_GENIC = TOTAL_SIM_GENIC/NUMBER,
        TOTAL_SIM_5UTR = sum(SIM_UTR5),
        PROP_SIM_5UTR = TOTAL_SIM_5UTR/NUMBER,
        TOTAL_SIM_3UTR = sum(SIM_UTR3),
        PROP_SIM_3UTR = TOTAL_SIM_3UTR/NUMBER,
        TOTAL_SIM_EXON = sum(SIM_EXON),
        PROP_SIM_EXON = TOTAL_SIM_EXON/NUMBER,
        TOTAL_SIM_INTRON = sum(SIM_INTRON),
        PROP_SIM_INTRON = TOTAL_SIM_INTRON/NUMBER,
        TOTAL_SIM_INTERGENIC = sum(SIM_INTERGENIC),
        PROP_SIM_INTERGENIC = TOTAL_SIM_INTERGENIC/NUMBER,
        TOTAL_SIM_UPSTREAM = sum(SIM_UPSTREAM),
        PROP_SIM_UPSTREAM = TOTAL_SIM_UPSTREAM/NUMBER,
        TOTAL_SIM_DOWNSTREAM = sum(SIM_DOWNSTREAM),
        PROP_SIM_DOWNSTREAM = TOTAL_SIM_DOWNSTREAM/NUMBER
      ) %>%
      ungroup()
  }
  
  sim_means_table <- rbindlist(sim_means)
  
  return(sim_means_table)
}
```

## Read in data
```{r}
sv <- fread("./Data/VCFs/SV_VCF/Parsed/parsed_ins_dels_pbsv.vcf")
#h1genes <- fread("./Data/GFFs/Parsed/hap1_genes_good.tsv")
#h2genes <- fread("./Data/GFFs/Parsed/hap2_genes_good.tsv")
h1gff <- fread("./Data/GFFs/Parsed/hap1_gff_good.tsv")
h2gff <- fread("./Data/GFFs/Parsed/hap2_gff_good.tsv")
combo_fasta <- ggread_fasta("./Data/Fastas/ref_chandler_combohap_default_scaffold.fasta")
fasta_5000 <- fread("./Data/Fastas/TEs/all_5000_fasta_table.tsv")
fasta_900 <- fread("./Data/Fastas/TEs/all_900_fasta_table.tsv")
```

## Combine gffs, create length column, make genes table
```{r}
gff <- rbind(h1gff, h2gff)
gff[, LENGTH := abs(STOP - START)]

genes <- gff[TYPE == "gene"]
```

## Add an upstream and downstream region to gff
```{r}
# Create upstream and downstream rows
upstream <- genes[, .(CHROM, SOURCE, TYPE = "upstream", 
                      START = pmax(1, START - 1000), 
                      STOP = START - 1, 
                      SCORE, DIRECTION, PHASE, INFO, GENE, 
                      UNIQUE = paste0("upstream_", UNIQUE), HAP, LENGTH)]

downstream <- genes[, .(CHROM, SOURCE, TYPE = "downstream", 
                        START = STOP + 1, 
                        STOP = STOP + 1000, 
                        SCORE, DIRECTION, PHASE, INFO, GENE, 
                        UNIQUE = paste0("downstream_", UNIQUE), HAP, LENGTH)]
stream <- rbind(upstream, downstream)

# Take strand into account
stream[DIRECTION == "-", TYPE := ifelse(TYPE == "upstream", "downstream", "upstream")]

gff <- rbind(gff, stream)

setorder(gff, CHROM, HAP, START)

gff
```


## Create new NAME column
```{r}
sv[, NUM := seq_len(.N), by = SVLEN]
sv[, NAME := paste(SVLEN, CHROM, POS, HAP, NUM, sep = "_")]
```

## Extract names vector
```{r}
fasta_5000[, NAME := gsub("_\\(reversed\\)", "", NAME)]
name_5000 <- fasta_5000$NAME

fasta_900[, NAME := gsub("_\\(reversed\\)", "", NAME)]
# Remove one of the split tandem duplicates
fasta_900[, NAME := sub("(*.)(-[0-9])", "\\1", NAME)]
fasta_900 <- fasta_900[-.N]
name_900 <- fasta_900$NAME
```

# Calculate distance of SVs to the nearest gene
## Label TEs
```{r}
sv[, TE := fifelse(NAME %in% name_5000, "te_5453", "sv")]
sv[, TE := fifelse(NAME %in% name_900, "te_899", TE)]

sv %>%
  group_by(SOURCE, TE) %>%
  summarise(
    NUMBER = n()
  )
```

## Find the nearest gene to each SV considering haplotype
```{r}
# Add columns
sv[, `:=`(CLOSEST_START = NA_integer_, CLOSEST_STOP = NA_integer_, GENIC = FALSE)]

# Create insertion only data
sv_ins <- sv[SVTYPE %in% "INS"]

# Loop to identify proximity to genes
in_near_gene(sv, genes)
in_near_gene(sv_ins, genes)
```

## Identify svs inside other genic features
```{r}
in_features(sv, gff)
in_features(sv_ins, gff)
```

## Summarize means of features
```{r}
real_means <- feature_means(sv)

real_means_ins <- feature_means(sv_ins)

setDT(real_means)
setDT(real_means_ins)
```

# Simulate te insertions
## Create fasta length table
```{r}
# Create table
fasta.table <- data.table(CHROM = names(combo_fasta), LENGTH = width(combo_fasta))
fasta.table <- fasta.table[grepl("NC_049", fasta.table$CHROM), ]

#Adjust and rename
fasta.table[, HAP := sub(".*(hap\\d+).*", "\\1", CHROM)]
fasta.table[, HAP := gsub("hap1", "A", HAP)]
fasta.table[, HAP := gsub("hap2", "B", HAP)]
fasta.table[, CHROM := as.numeric(sub("NC_0499(..).*", "\\1", CHROM))]
```

## Add a length column to sv
```{r}
sv <- merge(sv, fasta.table[, .(CHROM, HAP, CHROM_LENGTH = LENGTH)], by = c("CHROM", "HAP"), all.x = TRUE)
sv_ins <- merge(sv_ins, fasta.table[, .(CHROM, HAP, CHROM_LENGTH = LENGTH)], by = c("CHROM", "HAP"), all.x = TRUE)
```

## Running Simulation Loop
```{r, eval=FALSE}
random_sv_list <- sim_in_near_gene(sv, genes)
save(random_sv_list, file = "./Data/VCFs/SV_VCF/Parsed/random_locations_denovo_sv_list.RData")

random_ins_list <- sim_in_near_gene(sv_ins, genes)
save(random_ins_list, file = "./Data/VCFs/SV_VCF/Parsed/random_locations_denovo_insertion_list.RData")
```

```{r, eval=FALSE}
load("./Data/VCFs/SV_VCF/Parsed/random_locations_denovo_insertion_list.RData")

# Set cores
cores=8

num_cores <- cores
cl <- makeCluster(num_cores)
registerDoParallel(cl)

# Run function
system.time({
  random_ins_list_features <- foreach(sv = random_ins_list, .packages = "data.table") %dopar% {
    
    sv <- sv[, .(CHROM, HAP, POS, SIM_POS, SOURCE, TE, CLOSEST_START, CLOSEST_STOP, GENIC, SIM_START, SIM_STOP, SIM_GENIC)]  # Adjust based on what's needed
    
    sim_in_feature(sv, gff)
  }
})

# Stop parallel
stopCluster(cl)

save(random_ins_list_features, file = "./Data/VCFs/SV_VCF/Parsed/random_locations_denovo_features_ins_list.RData")
```

## Read in generated list
```{r}
load("./Data/VCFs/SV_VCF/Parsed/random_locations_denovo_features_ins_list.RData")
```

## Summarize list all features
```{r, message=FALSE}
sim_feat_means <- sim_feature_list_means(random_ins_list_features)
```

## Reshape data for genic proportion
```{r}
real_means_ins[, UNIQUE := paste(SOURCE, TE, sep = "_")]

sim_feat_means[, UNIQUE := paste(SOURCE, TE, sep = "_")]
```

## Identify pseudo-pvalues for features
```{r, eval=FALSE}
prop.table(table(sim_feat_means[TE %in% "te_899", MEAN_SIMSTART] < real_means_ins[TE %in% "te_899", MEAN_START]))
prop.table(table(sim_feat_means[TE %in% "te_899", MEAN_SIMSTOP] < real_means_ins[TE %in% "te_899", MEAN_STOP]))
prop.table(table(sim_feat_means[TE %in% "te_899", PROP_SIM_GENIC] < real_means_ins[TE %in% "te_899", PROP_GENIC]))
prop.table(table(sim_feat_means[TE %in% "te_899", PROP_SIM_5UTR] < real_means_ins[TE %in% "te_899", PROP_5UTR]))
prop.table(table(sim_feat_means[TE %in% "te_899", PROP_SIM_3UTR] < real_means_ins[TE %in% "te_899", PROP_3UTR]))
prop.table(table(sim_feat_means[TE %in% "te_899", PROP_SIM_EXON] < real_means_ins[TE %in% "te_899", PROP_EXON]))
prop.table(table(sim_feat_means[TE %in% "te_899", PROP_SIM_INTRON] < real_means_ins[TE %in% "te_899", PROP_INTRON]))
prop.table(table(sim_feat_means[TE %in% "te_899", PROP_SIM_INTERGENIC] < real_means_ins[TE %in% "te_899", PROP_INTERGENIC]))
prop.table(table(sim_feat_means[TE %in% "te_899", PROP_SIM_UPSTREAM] < real_means_ins[TE %in% "te_899", PROP_UPSTREAM]))
prop.table(table(sim_feat_means[TE %in% "te_899", PROP_SIM_DOWNSTREAM] < real_means_ins[TE %in% "te_899", PROP_DOWNSTREAM]))


prop.table(table(sim_feat_means[TE %in% "te_5453", MEAN_SIMSTART] < real_means_ins[TE %in% "te_5453", MEAN_START]))
prop.table(table(sim_feat_means[TE %in% "te_5453", MEAN_SIMSTOP] < real_means_ins[TE %in% "te_5453", MEAN_STOP]))
prop.table(table(sim_feat_means[TE %in% "te_5453", PROP_SIM_GENIC] < real_means_ins[TE %in% "te_5453", PROP_GENIC]))
prop.table(table(sim_feat_means[TE %in% "te_5453", PROP_SIM_5UTR] < real_means_ins[TE %in% "te_5453", PROP_5UTR]))
prop.table(table(sim_feat_means[TE %in% "te_5453", PROP_SIM_3UTR] < real_means_ins[TE %in% "te_5453", PROP_3UTR]))
prop.table(table(sim_feat_means[TE %in% "te_5453", PROP_SIM_EXON] < real_means_ins[TE %in% "te_5453", PROP_EXON]))
prop.table(table(sim_feat_means[TE %in% "te_5453", PROP_SIM_INTRON] < real_means_ins[TE %in% "te_5453", PROP_INTRON]))
prop.table(table(sim_feat_means[TE %in% "te_5453", PROP_SIM_INTERGENIC] < real_means_ins[TE %in% "te_5453", PROP_INTERGENIC]))
prop.table(table(sim_feat_means[TE %in% "te_5453", PROP_SIM_UPSTREAM] < real_means_ins[TE %in% "te_5453", PROP_UPSTREAM]))
prop.table(table(sim_feat_means[TE %in% "te_5453", PROP_SIM_DOWNSTREAM] < real_means_ins[TE %in% "te_5453", PROP_DOWNSTREAM]))
```

## Test out plotting other features TE 899
```{r}
sim_feat_means %>%
  filter(TE %in% "te_899") %>%
  ggplot(aes(x = PROP_SIM_GENIC, y = UNIQUE)) +
  geom_violin(aes(fill = SOURCE),
              alpha = 0.95,
              color = "grey70") +
  geom_point(data = real_means_ins[TE %in% "te_899"],
             aes(x = PROP_GENIC,
                 y = UNIQUE,
                 color = TE)) +
  geom_segment(aes(x = 0.2933971, xend = 0.1714286,  y = 1.41, yend = 1.41),
               linewidth = 0.175, color = "black") +
  geom_segment(aes(x = 0.2933971, xend = 0.2933971, y = 1.36, yend = 1.41),
               linewidth = 0.175, color = "black") +
  geom_segment(aes(x = 0.1714286, xend = 0.1714286, y = 1.36, yend = 1.41),
               linewidth = 0.175, color = "black") +
  annotate("text", x = (0.2933971 + 0.1714286) / 2, y = 1.525,
           label = "0.0326", size = 2, fontface = "bold", color = "black") +
  scale_fill_manual(
    values = c("cr2" = "grey90"
    ),
    labels = c("cr2" = "Simulation"
    )) +
  scale_color_manual(
    values = c(
      "te_899" = "#7287eb"
    ),
    labels = c(
      "te_899" = "Observed")) +
  scale_x_continuous(
    breaks = c(0, 0.3, 0.6),
    labels = c("0", "0.3", "0.6")
  ) +
  labs(title = "Genes", fill = "Simulated distribution", color = "Observed") +
  guides(fill = guide_legend(order = 1), color = guide_legend(order = 2)) +
  theme_classic(base_size = 6) +
  theme(    
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    axis.title.x = element_blank(),
    legend.title = element_blank(),
    #legend.position = c(0.81,0.85),
    legend.position = "none",
    legend.box = "veritcal",
    legend.box.spacing = unit(0, "cm"),
    legend.spacing.y = unit(-0.2, "cm"),
    legend.key.size = unit(0.2, "cm"),
    legend.background = element_rect(fill = NA, color = NA),
    legend.key = element_rect(fill = NA, color = NA),
    panel.background = element_rect(fill = NA, color = NA),
    plot.title = element_text(hjust = 0.5),
    plot.background = element_rect(fill = NA, color = NA),
    plot.margin = margin(t = 0.1, r = 0.2, b = 0.1, l = 0.01, unit = "cm")
  )
ggsave("./Figures/Transposable_Elements/ins_simulation_v_real_genes_900.pdf", height = 0.78,  width = 0.935)
#ggsave("./Figures/Transposable_Elements/ins_simulation_v_real_genes_900_label.pdf", height = 0.75, width = 1.3)

sim_feat_means %>%
  filter(TE %in% "te_899") %>%
  ggplot(aes(x = PROP_SIM_INTERGENIC, y = UNIQUE)) +
  geom_violin(aes(fill = SOURCE),
              alpha = 0.95,
              color = "grey70") +
  geom_point(data = real_means_ins[TE %in% "te_899"],
             aes(x = PROP_INTERGENIC,
                 y = UNIQUE,
                 color = TE)) +
  geom_segment(aes(x = 0.71396, xend = 0.8285714,  y = 1.41, yend = 1.41),
               linewidth = 0.175, color = "black") +
  geom_segment(aes(x = 0.71396, xend = 0.71396, y = 1.36, yend = 1.41),
               linewidth = 0.175, color = "black") +
  geom_segment(aes(x = 0.8285714, xend = 0.8285714, y = 1.36, yend = 1.41),
               linewidth = 0.175, color = "black") +
  annotate("text", x = (0.71396 + 0.8285714) / 2, y = 1.525,
           label = "ns", size = 2) +
  scale_fill_manual(
    values = c("cr2" = "grey90"
    ),
    labels = c("cr2" = "Simulation"
    )) +
  scale_color_manual(
    values = c(
      "te_899" = "#7287eb"
    ),
    labels = c(
      "te_899" = "900 class")) +
  scale_x_continuous(
    breaks = c(0.5, 0.7, 0.9)
    #labels = function(x) ifelse(x == 0, "0", scales::label_number()(x))
  ) +
  labs(title ="Intergenic", fill = "Simulated distribution", color = "Observed") +
  guides(fill = guide_legend(order = 1), color = guide_legend(order = 2)) +
  theme_classic(base_size = 6) +
  theme(    
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    legend.title = element_blank(),
    legend.position = "none",
    legend.box = "veritcal",
    legend.box.spacing = unit(0, "cm"),
    legend.spacing.y = unit(-0.2, "cm"),
    legend.key.size = unit(0.2, "cm"),
    legend.background = element_rect(fill = NA, color = NA),
    legend.key = element_rect(fill = NA, color = NA),
    panel.background = element_rect(fill = NA, color = NA),
    plot.background = element_rect(fill = NA, color = NA),
    plot.margin = margin(t = 0.1, r = 0.2, b = 0.1, l = 0.01, unit = "cm"),
    axis.title.x = element_blank(),
    plot.title = element_text(hjust = 0.5)
  )
ggsave("./Figures/Transposable_Elements/ins_simulation_v_real_intergenic_900.pdf", height = 0.78,  width = 0.935)

sim_feat_means %>%
  filter(TE %in% "te_899") %>%
  ggplot(aes(x = PROP_SIM_EXON, y = UNIQUE)) +
  geom_violin(aes(fill = SOURCE),
              alpha = 0.95,
              color = "grey70") +
  geom_point(data = real_means_ins[TE %in% "te_899"],
             aes(x = PROP_EXON,
                 y = UNIQUE,
                 color = TE)) +
  geom_segment(aes(x = 0.1031171, xend = 0.1142857,  y = 1.41, yend = 1.41),
               linewidth = 0.175, color = "black") +
  geom_segment(aes(x = 0.1031171, xend = 0.1031171, y = 1.36, yend = 1.41),
               linewidth = 0.175, color = "black") +
  geom_segment(aes(x = 0.1142857, xend = 0.1142857, y = 1.36, yend = 1.41),
               linewidth = 0.175, color = "black") +
  annotate("text", x = (0.1031171 + 0.1142857) / 2, y = 1.525,
           label = "ns", size = 2) +
  scale_fill_manual(
    values = c("cr2" = "grey90"
    ),
    labels = c("cr2" = "Simulation"
    )) +
  scale_color_manual(
    values = c(
      "te_899" = "#7287eb"
    ),
    labels = c(
      "te_899" = "900 class")) +
  scale_x_continuous(
    breaks = c(0, 0.15, 0.3),
    labels = c("0", "0.15", "0.3")
  ) +
  labs(title ="Exons", fill = "Simulated distribution", color = "Observed") +
  guides(fill = guide_legend(order = 1), color = guide_legend(order = 2)) +
  theme_classic(base_size = 6) +
  theme(    
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    legend.title = element_blank(),
    legend.position = "none",
    legend.box = "none",
    legend.box.spacing = unit(0, "cm"),
    legend.spacing.y = unit(-0.2, "cm"),
    legend.key.size = unit(0.2, "cm"),
    legend.background = element_rect(fill = NA, color = NA),
    legend.key = element_rect(fill = NA, color = NA),
    panel.background = element_rect(fill = NA, color = NA),
    plot.background = element_rect(fill = NA, color = NA),
    plot.margin = margin(t = 0.1, r = 0.2, b = 0.1, l = 0.01, unit = "cm"),
    axis.title.x = element_blank(),
    plot.title = element_text(hjust = 0.5)
  )
ggsave("./Figures/Transposable_Elements/ins_simulation_v_real_exon_900.pdf", height = 0.78,  width = 0.935)

sim_feat_means %>%
  filter(TE %in% "te_899") %>%
  ggplot(aes(x = PROP_SIM_INTRON, y = UNIQUE)) +
  geom_violin(aes(fill = SOURCE),
              alpha = 0.95,
              color = "grey70") +
  geom_point(data = real_means_ins[TE %in% "te_899"],
             aes(x = PROP_INTRON,
                 y = UNIQUE,
                 color = TE)) +
  geom_segment(aes(x = 0.1913143, xend = 0.05714286,  y = 1.41, yend = 1.41),
               linewidth = 0.175, color = "black") +
  geom_segment(aes(x = 0.1913143, xend = 0.1913143, y = 1.36, yend = 1.41),
               linewidth = 0.175, color = "black") +
  geom_segment(aes(x = 0.05714286, xend = 0.05714286, y = 1.36, yend = 1.41),
               linewidth = 0.175, color = "black") +
  annotate("text", x = (0.1913143 + 0.05714286) / 2, y = 1.525,
           label = "0.0058", size = 2, fontface = "bold", color = "black") +
  scale_fill_manual(
    values = c("cr2" = "grey90"
    ),
    labels = c("cr2" = "Simulation"
    )) +
  scale_color_manual(
    values = c(
      "te_899" = "#7287eb"
    ),
    labels = c(
      "te_899" = "900 class")) +
  scale_x_continuous(
    breaks = c(0, 0.2, 0.4)
    #labels = function(x) ifelse(x == 0, "0", scales::label_number()(x))
  ) +
  labs(title ="Introns", fill = "Simulated distribution", color = "Observed") +
  guides(fill = guide_legend(order = 1), color = guide_legend(order = 2)) +
  theme_classic(base_size = 6) +
  theme(    
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    legend.title = element_blank(),
    legend.position = "none",
    legend.box = "veritcal",
    legend.box.spacing = unit(0, "cm"),
    legend.spacing.y = unit(-0.2, "cm"),
    legend.key.size = unit(0.2, "cm"),
    legend.background = element_rect(fill = NA, color = NA),
    legend.key = element_rect(fill = NA, color = NA),
    panel.background = element_rect(fill = NA, color = NA),
    plot.background = element_rect(fill = NA, color = NA),
    plot.margin = margin(t = 0.1, r = 0.2, b = 0.1, l = 0.01, unit = "cm"),
    axis.title.x = element_blank(),
    plot.title = element_text(hjust = 0.5)
  )
ggsave("./Figures/Transposable_Elements/ins_simulation_v_real_intron_900.pdf", height = 0.78,  width = 0.935)

sim_feat_means %>%
  filter(TE %in% "te_899") %>%
  ggplot(aes(x = PROP_SIM_5UTR, y = UNIQUE)) +
  geom_violin(aes(fill = SOURCE),
              alpha = 0.95,
              color = "grey70") +
  geom_point(data = real_means_ins[TE %in% "te_899"],
             aes(x = PROP_5UTR,
                 y = UNIQUE,
                 color = TE)) +
  geom_segment(aes(x = 0.01146857, xend = 0.02857143,  y = 1.41, yend = 1.41),
               linewidth = 0.175, color = "black") +
  geom_segment(aes(x = 0.01146857, xend = 0.01146857, y = 1.36, yend = 1.41),
               linewidth = 0.175, color = "black") +
  geom_segment(aes(x = 0.02857143, xend = 0.02857143, y = 1.36, yend = 1.41),
               linewidth = 0.175, color = "black") +
  annotate("text", x = (0.01146857 + 0.02857143) / 2, y = 1.525,
           label = "ns", size = 2) +
  scale_fill_manual(
    values = c("cr2" = "grey90"
    ),
    labels = c("cr2" = "Simulation"
    )) +
  scale_color_manual(
    values = c(
      "te_899" = "#7287eb"
    ),
    labels = c(
      "te_899" = "900 class")) +
  scale_x_continuous(
    breaks = c(0, 0.05, .1),
    labels = c("0", "0.05", "0.1")
  ) +
  labs(title ="5' UTR", fill = "Simulated distribution", color = "Observed") +
  guides(fill = guide_legend(order = 1), color = guide_legend(order = 2)) +
  theme_classic(base_size = 6) +
  theme(    
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    legend.title = element_blank(),
    legend.position = "none",
    legend.box = "none",
    legend.box.spacing = unit(0, "cm"),
    legend.spacing.y = unit(-0.2, "cm"),
    legend.key.size = unit(0.2, "cm"),
    legend.background = element_rect(fill = NA, color = NA),
    legend.key = element_rect(fill = NA, color = NA),
    panel.background = element_rect(fill = NA, color = NA),
    plot.background = element_rect(fill = NA, color = NA),
    plot.margin = margin(t = 0.1, r = 0.2, b = 0.1, l = 0.01, unit = "cm"),
    axis.title.x = element_blank(),
    plot.title = element_text(hjust = 0.5)
  )
ggsave("./Figures/Transposable_Elements/ins_simulation_v_real_5utr_900.pdf", height = 0.78,  width = 0.935)

sim_feat_means %>%
  filter(TE %in% "te_899") %>%
  ggplot(aes(x = PROP_SIM_3UTR, y = UNIQUE)) +
  geom_violin(aes(fill = SOURCE),
              alpha = 0.95,
              color = "grey70") +
  geom_point(data = real_means_ins[TE %in% "te_899"],
             aes(x = PROP_3UTR,
                 y = UNIQUE,
                 color = TE)) +
  geom_segment(aes(x = 0.01089714, xend = 0.02857143,  y = 1.41, yend = 1.41),
               linewidth = 0.175, color = "black") +
  geom_segment(aes(x = 0.01089714, xend = 0.01089714, y = 1.36, yend = 1.41),
               linewidth = 0.175, color = "black") +
  geom_segment(aes(x = 0.02857143, xend = 0.02857143, y = 1.36, yend = 1.41),
               linewidth = 0.175, color = "black") +
  annotate("text", x = (0.01089714 + 0.02857143) / 2, y = 1.525,
           label = "ns", size = 2) +
  scale_fill_manual(
    values = c("cr2" = "grey90"
    ),
    labels = c("cr2" = "Simulation"
    )) +
  scale_color_manual(
    values = c(
      "te_899" = "#7287eb"
    ),
    labels = c(
      "te_899" = "900 class")) +
  scale_x_continuous(
    breaks = c(0, 0.05, .1),
    labels = c("0", "0.05", "0.1")
  ) +
  labs(title ="3' UTR", fill = "Simulated distribution", color = "Observed") +
  guides(fill = guide_legend(order = 1), color = guide_legend(order = 2)) +
  theme_classic(base_size = 6) +
  theme(    
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    legend.title = element_blank(),
    legend.position = "none",
    legend.box = "none",
    legend.box.spacing = unit(0, "cm"),
    legend.spacing.y = unit(-0.2, "cm"),
    legend.key.size = unit(0.2, "cm"),
    legend.background = element_rect(fill = NA, color = NA),
    legend.key = element_rect(fill = NA, color = NA),
    panel.background = element_rect(fill = NA, color = NA),
    plot.background = element_rect(fill = NA, color = NA),
    plot.margin = margin(t = 0.1, r = 0.2, b = 0.1, l = 0.01, unit = "cm"),
    axis.title.x = element_blank(),
    plot.title = element_text(hjust = 0.5)
  )
ggsave("./Figures/Transposable_Elements/ins_simulation_v_real_3utr_900.pdf", height = 0.78,  width = 0.935)

sim_feat_means %>%
  filter(TE %in% "te_899") %>%
  ggplot(aes(x = PROP_SIM_UPSTREAM, y = UNIQUE)) +
  geom_violin(aes(fill = SOURCE),
              alpha = 0.95,
              color = "grey70") +
  geom_point(data = real_means_ins[TE %in% "te_899"],
             aes(x = PROP_UPSTREAM,
                 y = UNIQUE,
                 color = TE)) +
  geom_segment(aes(x = 0.05456571, xend = 0.2285714, y = 1.41, yend = 1.41),
               linewidth = 0.175, color = "black") +
  geom_segment(aes(x = 0.05456571, xend = 0.05456571, y = 1.36, yend = 1.41),
               linewidth = 0.175, color = "black") +
  geom_segment(aes(x = 0.2285714, xend = 0.2285714, y = 1.36, yend = 1.41),
               linewidth = 0.175, color = "black") +
  annotate("text", x = (0.05456571 + 0.2285714) / 2, y = 1.525,
           label = "0.0006", size = 2, fontface = "bold", color = "black") +
  scale_fill_manual(
    values = c("cr2" = "grey90"
    ),
    labels = c("cr2" = "Simulation"
    )) +
  scale_color_manual(
    values = c(
      "te_899" = "#7287eb"
    ),
    labels = c(
      "te_899" = "900 class")) +
  scale_x_continuous(
    labels = function(x) ifelse(x == 0, "0", scales::label_number()(x))
  ) +
  labs(title ="Upstream", fill = "Simulated distribution", color = "Observed") +
  guides(fill = guide_legend(order = 1), color = guide_legend(order = 2)) +
  theme_classic(base_size = 6) +
  theme(    
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    legend.title = element_blank(),
    legend.position = "none",
    legend.box = "veritcal",
    legend.box.spacing = unit(0, "cm"),
    legend.spacing.y = unit(-0.2, "cm"),
    legend.key.size = unit(0.2, "cm"),
    legend.background = element_rect(fill = NA, color = NA),
    legend.key = element_rect(fill = NA, color = NA),
    panel.background = element_rect(fill = NA, color = NA),
    plot.background = element_rect(fill = NA, color = NA),
    plot.margin = margin(t = 0.1, r = 0.2, b = 0.1, l = 0.01, unit = "cm"),
    axis.title.x = element_blank(),
    plot.title = element_text(hjust = 0.5)
  )
ggsave("./Figures/Transposable_Elements/ins_simulation_v_real_upstream_900.pdf", height = 0.78,  width = 0.935)

sim_feat_means %>%
  filter(TE %in% "te_899") %>%
  ggplot(aes(x = PROP_SIM_DOWNSTREAM, y = UNIQUE)) +
  geom_violin(aes(fill = SOURCE),
              alpha = 0.95,
              color = "grey70") +
  geom_point(data = real_means_ins[TE %in% "te_899"],
             aes(x = PROP_DOWNSTREAM,
                 y = UNIQUE,
                 color = TE)) +
  geom_segment(aes(x = 0.05299714, xend = 0.1428571,  y = 1.41, yend = 1.41),
               linewidth = 0.175, color = "black") +
  geom_segment(aes(x = 0.05299714, xend = 0.05299714, y = 1.36, yend = 1.41),
               linewidth = 0.175, color = "black") +
  geom_segment(aes(x = 0.1428571, xend = 0.1428571, y = 1.36, yend = 1.41),
               linewidth = 0.175, color = "black") +
  annotate("text", x = (0.05299714 + 0.1428571) / 2, y = 1.525,
           label = "0.0385", size = 2, fontface = "bold", color = "black") +
  scale_fill_manual(
    values = c("cr2" = "grey90"
    ),
    labels = c("cr2" = "Simulation"
    )) +
  scale_color_manual(
    values = c(
      "te_899" = "#7287eb"
    ),
    labels = c(
      "te_899" = "900 class")) +
  scale_x_continuous(
    breaks = c(0, 0.1, .2),
    labels = function(x) ifelse(x == 0, "0", scales::label_number()(x))
  ) +
  labs(title ="Downstream", fill = "Simulated distribution", color = "Observed") +
  guides(fill = guide_legend(order = 1), color = guide_legend(order = 2)) +
  theme_classic(base_size = 6) +
  theme(    
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    legend.title = element_blank(),
    legend.position = "none",
    legend.box = "none",
    legend.box.spacing = unit(0, "cm"),
    legend.spacing.y = unit(-0.2, "cm"),
    legend.key.size = unit(0.2, "cm"),
    legend.background = element_rect(fill = NA, color = NA),
    legend.key = element_rect(fill = NA, color = NA),
    panel.background = element_rect(fill = NA, color = NA),
    plot.background = element_rect(fill = NA, color = NA),
    plot.margin = margin(t = 0.1, r = 0.2, b = 0.1, l = 0.01, unit = "cm"),
    axis.title.x = element_blank(),
    plot.title = element_text(hjust = 0.5)
  )
ggsave("./Figures/Transposable_Elements/ins_simulation_v_real_downstream_900.pdf", height = 0.78,  width = 0.935)

```

## Test out plotting other features TE 5453
```{r}
sim_feat_means %>%
  filter(TE %in% "te_5453") %>%
  ggplot(aes(x = PROP_SIM_GENIC, y = UNIQUE)) +
  geom_violin(aes(fill = SOURCE),
              alpha = 0.95,
              color = "grey70") +
  geom_point(data = real_means_ins[TE %in% "te_5453"],
             aes(x = PROP_GENIC,
                 y = UNIQUE,
                 color = TE)) +
  geom_segment(aes(x = 0.519084, xend = 0.2803687,  y = 1.41, yend = 1.41),
               linewidth = 0.175, color = "black") +
  geom_segment(aes(x = 0.519084, xend = 0.519084, y = 1.36, yend = 1.41),
               linewidth = 0.175, color = "black") +
  geom_segment(aes(x = 0.2803687, xend = 0.2803687, y = 1.36, yend = 1.41),
               linewidth = 0.175, color = "black") +
  annotate("text", x = (0.519084 + 0.2803687) / 2, y = 1.525,
           label = "0", size = 2, fontface = "bold", color = "black") +
  scale_fill_manual(
    values = c("cr2" = "grey90"
    ),
    labels = c("cr2" = "Simulation"
    )) +
  scale_color_manual(
    values = c(
      "te_5453" = "#eb7287"
    ),
    labels = c(
      "te_5453" = "Observed")) +
  scale_x_continuous(
    breaks = c(0.15, 0.3, 0.45),
    labels = c("0.15", "0.3", "0.45")
  ) +
  labs(title ="Genes", fill = "Simulated distribution", color = "Observed") +
  guides(fill = guide_legend(order = 1), color = guide_legend(order = 2)) +
  theme_classic(base_size = 6) +
  theme(    
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    legend.title = element_blank(),
    #legend.position = c(0.77,0.82),
    legend.position = "none",
    legend.box = "veritcal",
    legend.box.spacing = unit(0, "cm"),
    legend.spacing.y = unit(-0.2, "cm"),
    legend.key.size = unit(0.2, "cm"),
    legend.background = element_rect(fill = NA, color = NA),
    legend.key = element_rect(fill = NA, color = NA),
    panel.background = element_rect(fill = NA, color = NA),
    plot.background = element_rect(fill = NA, color = NA),
    plot.margin = margin(t = 0.1, r = 0.2, b = 0.1, l = 0.01, unit = "cm"),
    axis.title.x = element_blank(),
    plot.title = element_text(hjust = 0.5)
  )
ggsave("./Figures/Transposable_Elements/ins_simulation_v_real_genes_5500.pdf", height = 0.78,  width = 0.935)
#ggsave("./Figures/Transposable_Elements/ins_simulation_v_real_genes_5500_label.pdf", height = 0.75, width = 1.3)

sim_feat_means %>%
  filter(TE %in% "te_5453") %>%
  ggplot(aes(x = PROP_SIM_INTERGENIC, y = UNIQUE)) +
  geom_violin(aes(fill = SOURCE),
              alpha = 0.95,
              color = "grey70") +
  geom_point(data = real_means_ins[TE %in% "te_5453"],
             aes(x = PROP_INTERGENIC,
                 y = UNIQUE,
                 color = TE)) +
  geom_segment(aes(x = 0.7254672, xend = 0.4885496,  y = 1.41, yend = 1.41),
               linewidth = 0.175, color = "black") +
  geom_segment(aes(x = 0.7254672, xend = 0.7254672, y = 1.36, yend = 1.41),
               linewidth = 0.175, color = "black") +
  geom_segment(aes(x = 0.4885496, xend = 0.4885496, y = 1.36, yend = 1.41),
               linewidth = 0.175, color = "black") +
  annotate("text", x = (0.7254672 + 0.4885496) / 2, y = 1.525,
           label = "0", size = 2, fontface = "bold", color = "black") +
  scale_fill_manual(
    values = c("cr2" = "grey90"
    ),
    labels = c("cr2" = "Simulation"
    )) +
  scale_color_manual(
    values = c(
      "te_5453" = "#eb7287"
    ),
    labels = c(
      "te_5453" = "5500 class")) +
  scale_x_continuous(
    breaks = c(0.55, 0.7, 0.85),
    labels = c("0.55", "0.7", "0.85")
  ) +
  labs(title ="Intergenic", fill = "Simulated distribution", color = "Observed") +
  guides(fill = guide_legend(order = 1), color = guide_legend(order = 2)) +
  theme_classic(base_size = 6) +
  theme(    
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    legend.title = element_blank(),
    legend.position = "none",
    legend.box = "veritcal",
    legend.box.spacing = unit(0, "cm"),
    legend.spacing.y = unit(-0.2, "cm"),
    legend.key.size = unit(0.2, "cm"),
    legend.background = element_rect(fill = NA, color = NA),
    legend.key = element_rect(fill = NA, color = NA),
    panel.background = element_rect(fill = NA, color = NA),
    plot.background = element_rect(fill = NA, color = NA),
    plot.margin = margin(t = 0.1, r = 0.2, b = 0.1, l = 0.01, unit = "cm"),
    axis.title.x = element_blank(),
    plot.title = element_text(hjust = 0.5)
  )
ggsave("./Figures/Transposable_Elements/ins_simulation_v_real_intergenic_5500.pdf", height = 0.78,  width = 0.935)

sim_feat_means %>%
  filter(TE %in% "te_5453") %>%
  ggplot(aes(x = PROP_SIM_EXON, y = UNIQUE)) +
  geom_violin(aes(fill = SOURCE),
              alpha = 0.95,
              color = "grey70") +
  geom_point(data = real_means_ins[TE %in% "te_5453"],
             aes(x = PROP_EXON,
                 y = UNIQUE,
                 color = TE)) +
  geom_segment(aes(x = 0.09857099, xend = 0.3587786,  y = 1.41, yend = 1.41),
               linewidth = 0.175, color = "black") +
  geom_segment(aes(x = 0.09857099, xend = 0.09857099, y = 1.36, yend = 1.41),
               linewidth = 0.175, color = "black") +
  geom_segment(aes(x = 0.3587786, xend = 0.3587786, y = 1.36, yend = 1.41),
               linewidth = 0.175, color = "black") +
  annotate("text", x = (0.09857099 + 0.3587786) / 2, y = 1.525,
           label = "0", size = 2, fontface = "bold", color = "black") +
  scale_fill_manual(
    values = c("cr2" = "grey90"
    ),
    labels = c("cr2" = "Simulation"
    )) +
  scale_color_manual(
    values = c(
      "te_5453" = "#eb7287"
    ),
    labels = c(
      "te_5453" = "5500 class")) +
  scale_x_continuous(
    breaks = c(0.05, 0.2, 0.35),
    labels = c("0.05", "0.2", "0.35")
  ) +
  labs(title ="Exons", fill = "Simulated distribution", color = "Observed") +
  guides(fill = guide_legend(order = 1), color = guide_legend(order = 2)) +
  theme_classic(base_size = 6) +
  theme(    
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    legend.title = element_blank(),
    legend.position = "none",
    legend.box = "none",
    legend.box.spacing = unit(0, "cm"),
    legend.spacing.y = unit(-0.2, "cm"),
    legend.key.size = unit(0.2, "cm"),
    legend.background = element_rect(fill = NA, color = NA),
    legend.key = element_rect(fill = NA, color = NA),
    panel.background = element_rect(fill = NA, color = NA),
    plot.background = element_rect(fill = NA, color = NA),
    plot.margin = margin(t = 0.1, r = 0.2, b = 0.1, l = 0.01, unit = "cm"),
    axis.title.x = element_blank(),
    plot.title = element_text(hjust = 0.5)
  )
ggsave("./Figures/Transposable_Elements/ins_simulation_v_real_exon_5500.pdf", height = 0.78,  width = 0.935)

sim_feat_means %>%
  filter(TE %in% "te_5453") %>%
  ggplot(aes(x = PROP_SIM_INTRON, y = UNIQUE)) +
  geom_violin(aes(fill = SOURCE),
              alpha = 0.95,
              color = "grey70") +
  geom_point(data = real_means_ins[TE %in% "te_5453"],
             aes(x = PROP_INTRON,
                 y = UNIQUE,
                 color = TE)) +
  geom_segment(aes(x = 0.1828603, xend = 0.1679389,  y = 1.41, yend = 1.41),
               linewidth = 0.175, color = "black") +
  geom_segment(aes(x = 0.1828603, xend = 0.1828603, y = 1.36, yend = 1.41),
               linewidth = 0.175, color = "black") +
  geom_segment(aes(x = 0.1679389, xend = 0.1679389, y = 1.36, yend = 1.41),
               linewidth = 0.175, color = "black") +
  annotate("text", x = (0.1828603 + 0.1679389) / 2, y = 1.525,
           label = "ns", size = 2) +
  scale_fill_manual(
    values = c("cr2" = "grey90"
    ),
    labels = c("cr2" = "Simulation"
    )) +
  scale_color_manual(
    values = c(
      "te_5453" = "#eb7287"
    ),
    labels = c(
      "te_5453" = "5500 class")) +
  scale_x_continuous(
    breaks = c(0.1, 0.2, 0.3),
    labels = c("0.1", "0.2", "0.3")
  ) +
  labs(title ="Introns", fill = "Simulated distribution", color = "Observed") +
  guides(fill = guide_legend(order = 1), color = guide_legend(order = 2)) +
  theme_classic(base_size = 6) +
  theme(    
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    legend.title = element_blank(),
    legend.position = "none",
    legend.box = "veritcal",
    legend.box.spacing = unit(0, "cm"),
    legend.spacing.y = unit(-0.2, "cm"),
    legend.key.size = unit(0.2, "cm"),
    legend.background = element_rect(fill = NA, color = NA),
    legend.key = element_rect(fill = NA, color = NA),
    panel.background = element_rect(fill = NA, color = NA),
    plot.background = element_rect(fill = NA, color = NA),
    plot.margin = margin(t = 0.1, r = 0.2, b = 0.1, l = 0.01, unit = "cm"),
    axis.title.x = element_blank(),
    plot.title = element_text(hjust = 0.5)
  )
ggsave("./Figures/Transposable_Elements/ins_simulation_v_real_intron_5500.pdf", height = 0.78,  width = 0.935)

sim_feat_means %>%
  filter(TE %in% "te_5453") %>%
  ggplot(aes(x = PROP_SIM_5UTR, y = UNIQUE)) +
  geom_violin(aes(fill = SOURCE),
              alpha = 0.95,
              color = "grey70") +
  geom_point(data = real_means_ins[TE %in% "te_5453"],
             aes(x = PROP_5UTR,
                 y = UNIQUE,
                 color = TE)) +
  geom_segment(aes(x = 0.01057481, xend = 0.01526718,  y = 1.41, yend = 1.41),
               linewidth = 0.175, color = "black") +
  geom_segment(aes(x = 0.01057481, xend = 0.01057481, y = 1.36, yend = 1.41),
               linewidth = 0.175, color = "black") +
  geom_segment(aes(x = 0.01526718, xend = 0.01526718, y = 1.36, yend = 1.41),
               linewidth = 0.175, color = "black") +
  annotate("text", x = (0.01057481 + 0.01526718) / 2, y = 1.525,
           label = "ns", size = 2) +
  # geom_segment(aes(x = -Inf, xend = Inf, y = Inf, yend = Inf), color = "black", linewidth = 0.5) +
  # geom_segment(aes(x = Inf, xend = Inf, y = -Inf, yend = Inf), color = "black", linewidth = 0.5) +
  # geom_segment(aes(x = -Inf, xend = Inf, y = -Inf, yend = -Inf), color = "black", linewidth = 0.5) +
  scale_fill_manual(
    values = c("cr2" = "grey90"
    ),
    labels = c("cr2" = "Simulation"
    )) +
  scale_color_manual(
    values = c(
      "te_5453" = "#eb7287"
    ),
    labels = c(
      "te_5453" = "5500 class")) +
  scale_x_continuous(
    breaks = c(0.01, 0.035, 0.06),
    labels = c("0.01", "0.035", "0.06")
  ) +
  labs(title ="5' UTR", fill = "Simulated distribution", color = "Observed") +
  guides(fill = guide_legend(order = 1), color = guide_legend(order = 2)) +
  theme_classic(base_size = 6) +
  theme(    
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    legend.title = element_blank(),
    legend.position = "none",
    legend.box = "none",
    legend.box.spacing = unit(0, "cm"),
    legend.spacing.y = unit(-0.2, "cm"),
    legend.key.size = unit(0.2, "cm"),
    legend.background = element_rect(fill = NA, color = NA),
    legend.key = element_rect(fill = NA, color = NA),
    panel.background = element_rect(fill = NA, color = NA),
    plot.background = element_rect(fill = NA, color = NA),
    plot.margin = margin(t = 0.1, r = 0.2, b = 0.1, l = 0.01, unit = "cm"),
    axis.title.x = element_blank(),
    plot.title = element_text(hjust = 0.5)
  )
ggsave("./Figures/Transposable_Elements/ins_simulation_v_real_5utr_5500.pdf", height = 0.78,  width = 0.935)

sim_feat_means %>%
  filter(TE %in% "te_5453") %>%
  ggplot(aes(x = PROP_SIM_3UTR, y = UNIQUE)) +
  geom_violin(aes(fill = SOURCE),
              alpha = 0.95,
              color = "grey70") +
  geom_point(data = real_means_ins[TE %in% "te_5453"],
             aes(x = PROP_3UTR,
                 y = UNIQUE,
                 color = TE)) +
  geom_segment(aes(x = 0.01007328, xend = 0.01526718,  y = 1.41, yend = 1.41),
               linewidth = 0.175, color = "black") +
  geom_segment(aes(x = 0.01007328, xend = 0.01007328, y = 1.36, yend = 1.41),
               linewidth = 0.175, color = "black") +
  geom_segment(aes(x = 0.01526718, xend = 0.01526718, y = 1.36, yend = 1.41),
               linewidth = 0.175, color = "black") +
  annotate("text", x = (0.01007328 + 0.01526718) / 2, y = 1.525,
           label = "ns", size = 2) +
  scale_fill_manual(
    values = c("cr2" = "grey90"
    ),
    labels = c("cr2" = "Simulation"
    )) +
  scale_color_manual(
    values = c(
      "te_5453" = "#eb7287"
    ),
    labels = c(
      "te_5453" = "5500 class")) +
  scale_x_continuous(
    breaks = c(0.01, 0.035, 0.06),
    labels = c("0.01", "0.035", "0.06")
  ) +
  labs(title ="3' UTR", fill = "Simulated distribution", color = "Observed") +
  guides(fill = guide_legend(order = 1), color = guide_legend(order = 2)) +
  theme_classic(base_size = 6) +
  theme(    
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    legend.title = element_blank(),
    legend.position = "none",
    legend.box = "none",
    legend.box.spacing = unit(0, "cm"),
    legend.spacing.y = unit(-0.2, "cm"),
    legend.key.size = unit(0.2, "cm"),
    legend.background = element_rect(fill = NA, color = NA),
    legend.key = element_rect(fill = NA, color = NA),
    panel.background = element_rect(fill = NA, color = NA),
    plot.background = element_rect(fill = NA, color = NA),
    plot.margin = margin(t = 0.1, r = 0.2, b = 0.1, l = 0.01, unit = "cm"),
    axis.title.x = element_blank(),
    plot.title = element_text(hjust = 0.5)
  )
ggsave("./Figures/Transposable_Elements/ins_simulation_v_real_3utr_5500.pdf", height = 0.78,  width = 0.935)

sim_feat_means %>%
  filter(TE %in% "te_5453") %>%
  ggplot(aes(x = PROP_SIM_UPSTREAM, y = UNIQUE)) +
  geom_violin(aes(fill = SOURCE),
              alpha = 0.95,
              color = "grey70") +
  geom_point(data = real_means_ins[TE %in% "te_5453"],
             aes(x = PROP_UPSTREAM,
                 y = UNIQUE,
                 color = TE)) +
  geom_segment(aes(x = 0.05241221, xend = 0.08396947,  y = 1.41, yend = 1.41),
               linewidth = 0.175, color = "black") +
  geom_segment(aes(x = 0.05241221, xend = 0.05241221, y = 1.36, yend = 1.41),
               linewidth = 0.175, color = "black") +
  geom_segment(aes(x = 0.08396947, xend = 0.08396947, y = 1.36, yend = 1.41),
               linewidth = 0.175, color = "black") +
  annotate("text", x = (0.05241221 + 0.08396947) / 2, y = 1.525,
           label = "ns", size = 2) +
  # geom_segment(aes(x = -Inf, xend = Inf, y = Inf, yend = Inf), color = "black", linewidth = 0.5) +
  # geom_segment(aes(x = Inf, xend = Inf, y = -Inf, yend = Inf), color = "black", linewidth = 0.5) +
  # geom_segment(aes(x = -Inf, xend = Inf, y = -Inf, yend = -Inf), color = "black", linewidth = 0.5) +
  scale_fill_manual(
    values = c("cr2" = "grey90"
    ),
    labels = c("cr2" = "Simulation"
    )) +
  scale_color_manual(
    values = c(
      "te_5453" = "#eb7287"
    ),
    labels = c(
      "te_5453" = "5500 class")) +
  scale_x_continuous(
    breaks = c(0.025, 0.075, 0.125),
    labels = c("0.025", "0.075", "0.125")
    ) +
  labs(title ="Upstream", fill = "Simulated distribution", color = "Observed") +
  guides(fill = guide_legend(order = 1), color = guide_legend(order = 2)) +
  theme_classic(base_size = 6) +
  theme(    
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    legend.title = element_blank(),
    legend.position = "none",
    legend.box = "veritcal",
    legend.box.spacing = unit(0, "cm"),
    legend.spacing.y = unit(-0.2, "cm"),
    legend.key.size = unit(0.2, "cm"),
    legend.background = element_rect(fill = NA, color = NA),
    legend.key = element_rect(fill = NA, color = NA),
    panel.background = element_rect(fill = NA, color = NA),
    plot.background = element_rect(fill = NA, color = NA),
    plot.margin = margin(t = 0.1, r = 0.2, b = 0.1, l = 0.01, unit = "cm"),
    axis.title.x = element_blank(),
    plot.title = element_text(hjust = 0.5)
  )
ggsave("./Figures/Transposable_Elements/ins_simulation_v_real_upstream_5500.pdf", height = 0.78,  width = 0.935)

sim_feat_means %>%
  filter(TE %in% "te_5453") %>%
  ggplot(aes(x = PROP_SIM_DOWNSTREAM, y = UNIQUE)) +
  geom_violin(aes(fill = SOURCE),
              alpha = 0.95,
              color = "grey70") +
  geom_point(data = real_means_ins[TE %in% "te_5453"],
             aes(x = PROP_DOWNSTREAM,
                 y = UNIQUE,
                 color = TE)) +
  geom_segment(aes(x = 0.05043817, xend = 0.1068702,  y = 1.41, yend = 1.41),
               linewidth = 0.175, color = "black") +
  geom_segment(aes(x = 0.05043817, xend = 0.05043817, y = 1.36, yend = 1.41),
               linewidth = 0.175, color = "black") +
  geom_segment(aes(x = 0.1068702, xend = 0.1068702, y = 1.36, yend = 1.41),
               linewidth = 0.175, color = "black") +
  annotate("text", x = (0.05043817 + 0.1068702) / 2, y = 1.525,
           label = "0.0069", size = 2, fontface = "bold", color = "black") +
  scale_fill_manual(
    values = c("cr2" = "grey90"
    ),
    labels = c("cr2" = "Simulation"
    )) +
  scale_color_manual(
    values = c(
      "te_5453" = "#eb7287"
    ),
    labels = c(
      "te_5453" = "5500 class")) +
  scale_x_continuous(
    breaks = c(0.025, 0.075, 0.125),
    labels = c("0.025", "0.075", "0.125")
  ) +
  labs(title ="Downstream", fill = "Simulated distribution", color = "Observed") +
  guides(fill = guide_legend(order = 1), color = guide_legend(order = 2)) +
  theme_classic(base_size = 6) +
  theme(    
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    legend.title = element_blank(),
    legend.position = "none",
    legend.box = "none",
    legend.box.spacing = unit(0, "cm"),
    legend.spacing.y = unit(-0.2, "cm"),
    legend.key.size = unit(0.2, "cm"),
    legend.background = element_rect(fill = NA, color = NA),
    legend.key = element_rect(fill = NA, color = NA),
    panel.background = element_rect(fill = NA, color = NA),
    plot.background = element_rect(fill = NA, color = NA),
    plot.margin = margin(t = 0.1, r = 0.2, b = 0.1, l = 0.01, unit = "cm"),
    axis.title.x = element_blank(),
    plot.title = element_text(hjust = 0.5)
  )
ggsave("./Figures/Transposable_Elements/ins_simulation_v_real_downstream_5500.pdf", height = 0.81, width = 0.93)
```

# Archive
## Plot distribution of data
```{r, eval=FALSE}
sv %>%
  ggplot(aes(x = CLOSEST_START, fill = TE)) +
  geom_histogram(alpha = 0.33)+
  facet_grid(~SOURCE) +
  theme_classic()

sv %>%
  ggplot(aes(x = CLOSEST_STOP, fill = TE)) +
  geom_histogram(alpha = 0.33)+
  facet_grid(~SOURCE) +
  theme_classic()

sv_ins %>%
  ggplot(aes(x = CLOSEST_START, fill = TE)) +
  geom_histogram(alpha = 0.33)+
  facet_grid(~SOURCE) +
  theme_classic()

sv_ins %>%
  ggplot(aes(x = CLOSEST_STOP, fill = TE)) +
  geom_histogram(alpha = 0.33)+
  facet_grid(~SOURCE) +
  theme_classic()
```
## Read in list and adjust
```{r, eval=FALSE}
load("./Data/VCFs/SV_VCF/Parsed/random_locations_denovo_sv_list.RData")

sim_sv_table <- rbindlist(random_sv_list)
sim_sv_table[, DATA := "simulated"]
head(sim_sv_table)

sim_ins_table <- rbindlist(random_ins_list)
sim_ins_table[, DATA := "simulated"]
head(sim_ins_table)
```

## Summarize each table
```{r, message=FALSE, eval=FALSE}
sim_sv_means <- sim_list_means(random_sv_list)
sim_ins_means <- sim_list_means(random_ins_list)
```

# Compare simulations to real data
## Create a combinded data table
```{r, eval=FALSE}
# sv[, DATA := "real"]
# sv_ins[, DATA := "real"]
# 
# small_sv <- sv[, .(CHROM, HAP, POS, UNIQUE, SVTYPE, SVLEN, VAF, SOURCE, TE, CLOSEST_START, CLOSEST_STOP, GENIC, DATA)]
# small_ins <- sv[, .(CHROM, HAP, POS, UNIQUE, SVTYPE, SVLEN, VAF, SOURCE, TE, CLOSEST_START, CLOSEST_STOP, GENIC, DATA)]
# 
# 
# small_sv_sim <- sim_sv_table[, .(CHROM, HAP, POS, UNIQUE, SVTYPE, SVLEN, VAF, SOURCE, TE, SIM_START, SIM_STOP, SIM_GENIC, DATA)]
# small_ins_sim <- sim_ins_table[, .(CHROM, HAP, POS, UNIQUE, SVTYPE, SVLEN, VAF, SOURCE, TE, SIM_START, SIM_STOP, SIM_GENIC, DATA)]
# 
# setnames(small_sv, c("CLOSEST_START", "CLOSEST_STOP"), c("START", "STOP"))
# setnames(small_ins, c("CLOSEST_START", "CLOSEST_STOP"), c("START", "STOP"))
# 
# setnames(small_sv_sim, c("SIM_START", "SIM_STOP", "SIM_GENIC"), c("START", "STOP", "GENIC"))
# setnames(small_ins_sim, c("SIM_START", "SIM_STOP", "SIM_GENIC"), c("START", "STOP", "GENIC"))
# 
# 
# combo_sv <- rbind(small_sv, small_sv_sim)
# combo_ins <- rbind(small_ins, small_ins_sim)
```

## Boxplot comparing START positions
```{r, eval=FALSE}
# combo_sv %>%
#   ggplot(aes(x = DATA, y = START)) +
#   geom_boxplot() +
#   facet_grid(SOURCE ~ TE) +
#   stat_compare_means(method = "wilcox.test", comparisons = list(c("real", "simulated"))) +
#   theme_classic()
# 
# combo_sv %>%
#   ggplot(aes(x = DATA, y = log(START))) +
#   geom_boxplot() +
#   facet_grid(SOURCE ~ TE) +
#   theme_classic()
# 
# combo_ins %>%
#   ggplot(aes(x = DATA, y = START)) +
#   geom_boxplot() +
#   facet_grid(SOURCE ~ TE) +
#   stat_compare_means(method = "wilcox.test", comparisons = list(c("real", "simulated"))) +
#   theme_classic()
# 
# combo_ins %>%
#   ggplot(aes(x = DATA, y = log(START))) +
#   geom_boxplot() +
#   facet_grid(SOURCE ~ TE) +
#   theme_classic()
```

## Boxplots comparing STOP positions
```{r, eval=FALSE}
# combo_sv %>%
#   ggplot(aes(x = DATA, y = STOP)) +
#   geom_boxplot() +
#   facet_grid(SOURCE ~ TE) +
#   stat_compare_means(method = "wilcox.test", comparisons = list(c("real", "simulated"))) +
#   theme_classic()
# 
# combo_sv %>%
#   ggplot(aes(x = DATA, y = log(STOP))) +
#   geom_boxplot() +
#   facet_grid(SOURCE ~ TE) +
#   theme_classic()
# 
# combo_ins %>%
#   ggplot(aes(x = DATA, y = STOP)) +
#   geom_boxplot() +
#   facet_grid(SOURCE ~ TE) +
#   stat_compare_means(method = "wilcox.test", comparisons = list(c("real", "simulated"))) +
#   theme_classic()
# 
# combo_ins %>%
#   ggplot(aes(x = DATA, y = log(STOP))) +
#   geom_boxplot() +
#   facet_grid(SOURCE ~ TE) +
#   theme_classic()
```

# Plot to compare data means
## Distribution of start proximity
```{r, eval=FALSE}
# real_means
# real_means_ins
# 
# sim_sv_means %>%
#   filter(TE == "te_5453") %>%
#   ggplot(aes(x = MEAN_SIMSTART)) +
#   geom_histogram() +
#   geom_vline(xintercept = 4389.588, linetype = 2, color = "red2") +
#   labs(x = "Distance from gene start (bp)", y = "Number of simulations") +
#   theme_classic() +
#   theme()
# 
# sim_ins_means %>%
#   filter(TE == "te_5453") %>%
#   ggplot(aes(x = MEAN_SIMSTART)) +
#   geom_histogram() +
#   geom_vline(xintercept = 4389.588, linetype = 2, color = "red2") +
#   labs(x = "Distance from gene start (bp)", y = "Number of simulations") +
#   theme_classic() +
#   theme()
```

## Distribution of stop proximity
```{r, eval=FALSE}
# real_means
# real_means_ins
# 
# sim_sv_means %>%
#   filter(TE == "te_5453") %>%
#   ggplot(aes(x = MEAN_SIMSTOP)) +
#   geom_histogram() +
#   geom_vline(xintercept = 4614.191, linetype = 2, color = "red2") +
#   labs(x = "Distance from gene start (bp)", y = "Number of simulations") +
#   theme_classic() +
#   theme()
# 
# sim_ins_means %>%
#   filter(TE == "te_5453") %>%
#   ggplot(aes(x = MEAN_SIMSTOP)) +
#   geom_histogram() +
#   geom_vline(xintercept = 4614.191, linetype = 2, color = "red2") +
#   labs(x = "Distance from gene start (bp)", y = "Number of simulations") +
#   theme_classic() +
#   theme()
```

## Comparing proportion of gene landing
```{r, eval=FALSE}
# real_means
# real_means_ins
# 
# sim_sv_means %>%
#   filter(TE == "te_5453") %>%
#   ggplot(aes(x = PROP_SIM_GENIC)) +
#   geom_histogram() +
#   geom_vline(xintercept = 0.5190840, linetype = 2, color = "red2") +
#   labs(x = "Proportion in gene bodies", y = "Number of simulations") +
#   theme_classic() +
#   theme()
# 
# sim_ins_means %>%
#   filter(TE == "te_5453") %>%
#   ggplot(aes(x = PROP_SIM_GENIC)) +
#   geom_histogram() +
#   geom_vline(xintercept = 0.5190840, linetype = 2, color = "red2") +
#   labs(x = "Proportion in gene bodies", y = "Number of simulations") +
#   theme_classic() +
#   theme()
```

## Violin plots
```{r, eval=FALSE}
# sim_sv_means %>%
#   filter(TE == "te_5453") %>%
#   ggplot(aes(x = MEAN_SIMSTART, y = TE)) +
#   geom_jitter(alpha = 0.1) +
#   geom_violin(fill = NA) +
#   geom_point(data = real_means[TE == "te_5453"], aes(x = MEAN_CLOSESTART, y = TE, color = TE)) +
#   scale_fill_manual(values = ) +
#   theme_classic() +
#   theme()
```

# Plot realtionships
## Reshape data to long for proximity to genes
```{r, eval=FALSE}
# prox_sim_sv_means <- sim_prox_table(sim_sv_means)
# prox_sim_ins_means <- sim_prox_table(sim_ins_means)
# 
# prox_real_means <- real_prox_table(real_means)
# prox_real_means_ins <- real_prox_table(real_means_ins)
```

## Reshape all features to long
```{r, eval=FALSE}
# prox_feat_means <- sim_feat_means %>%
#   pivot_longer(cols = c(MEAN_SIMSTART, MEAN_SIMSTOP), names_to = "LOCATION", values_to = "DISTANCE")
# setDT(prox_feat_means)
# prox_feat_means[, LOCATION := fifelse(LOCATION == "MEAN_SIMSTART", "start", "stop")]
# prox_feat_means[, UNIQUE := paste(SOURCE, TE, LOCATION, sep = "_")]
```

## Reshape data for genic proportion
```{r, eval=FALSE}
sim_sv_means[, UNIQUE := paste(SOURCE, TE, sep = "_")]
sim_sv_means[, GENIC := PROP_SIM_GENIC]

sim_ins_means[, UNIQUE := paste(SOURCE, TE, sep = "_")]
sim_ins_means[, GENIC := PROP_SIM_GENIC]

real_means[, UNIQUE := paste(SOURCE, TE, sep = "_")]
real_means[, GENIC := PROP_GENIC]

real_means_ins[, UNIQUE := paste(SOURCE, TE, sep = "_")]
real_means_ins[, GENIC := PROP_GENIC]

sim_feat_means[, UNIQUE := paste(SOURCE, TE, sep = "_")]
```

## Identify pseudo-pvalues
```{r, eval=FALSE}
prop.table(table(sim_ins_means[TE %in% "te_899", MEAN_SIMSTART] < real_means_ins[TE %in% "te_899", MEAN_START]))
prop.table(table(sim_ins_means[TE %in% "te_899", MEAN_SIMSTOP] < real_means_ins[TE %in% "te_899", MEAN_STOP]))
prop.table(table(sim_ins_means[TE %in% "te_899", PROP_SIM_GENIC] < real_means_ins[TE %in% "te_899", PROP_GENIC]))

prop.table(table(sim_ins_means[TE %in% "te_5453", MEAN_SIMSTART] < real_means_ins[TE %in% "te_5453", MEAN_CLOSESTART]))
prop.table(table(sim_ins_means[TE %in% "te_5453", MEAN_SIMSTOP] < real_means_ins[TE %in% "te_5453", MEAN_CLOSESTOP]))
prop.table(table(sim_ins_means[TE %in% "te_5453", PROP_SIM_GENIC] < real_means_ins[TE %in% "te_5453", PROP_GENIC]))
```

## Setting levels
```{r, eval=FALSE}
means_levels <- c("ref_chandler_sv_stop", "ref_chandler_sv_start", "term_chandler_sv_stop", "term_chandler_sv_start", "cr2_te_5453_stop", "cr2_te_5453_start","cr2_te_899_stop","cr2_te_899_start", "cr2_sv_stop", "cr2_sv_start")
genes_levels <- c("ref_chandler_sv", "term_chandler_sv", "cr2_te_5453", "cr2_te_899", "cr2_sv")

prox_sim_sv_means$UNIQUE <- factor(prox_sim_sv_means$UNIQUE, levels = means_levels)
prox_sim_ins_means$UNIQUE <- factor(prox_sim_ins_means$UNIQUE, levels = means_levels)

prox_real_means$UNIQUE <- factor(prox_real_means$UNIQUE, levels = means_levels)
prox_real_means_ins$UNIQUE <- factor(prox_real_means_ins$UNIQUE, levels = means_levels)


sim_sv_means$UNIQUE <- factor(sim_sv_means$UNIQUE, levels = genes_levels)
sim_ins_means$UNIQUE <- factor(sim_ins_means$UNIQUE, levels = genes_levels)

real_means$UNIQUE <- factor(real_means$UNIQUE, levels = genes_levels)
real_means_ins$UNIQUE <- factor(real_means_ins$UNIQUE, levels = genes_levels)
```

## Plot proximity to both start and stop
```{r, eval=FALSE}
prox_sim_sv_means %>%
  ggplot(aes(x = DISTANCE, y = UNIQUE)) +
  geom_violin(aes(fill = SOURCE), alpha = 0.75) +
  geom_point(data = prox_real_means, aes(x = DISTANCE, y = UNIQUE, color = TE)) +
  scale_fill_manual(
    values = c("cr2" = "goldenrod1",
               "ref_chandler" = "darkseagreen4",
               "term_chandler" = "darkseagreen3"),
    labels = c("cr2" = "E HiFi",
               "ref_chandler" = "Reference",
               "term_chandler" = "Tree")) +
  scale_color_manual(
    values = c("sv" = "grey30",
               "te_5453" = "#DA8392",
               "te_899" = "#FBC5AD"),
    labels = c("sv" = "Other SV",
               "te_5453" = "5453 bp TE",
               "te_899" = "899bp TE")) +
  labs(x = "Distance from site (bp)") +
  theme_classic(base_size = 6) +
  theme(
    axis.title.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.key.size = unit(0.2, "cm"),
    legend.position = c(0.8,0.9)
  )
```

## Plot proximity to start
```{r, eval=FALSE}
prox_sim_sv_means %>%
  filter(LOCATION == "start") %>%
  ggplot(aes(x = DISTANCE, y = UNIQUE)) +
  geom_violin(aes(fill = SOURCE), alpha = 0.75) +
  geom_point(data = prox_real_means[LOCATION == "start"], aes(x = DISTANCE, y = UNIQUE, color = TE)) +
  scale_fill_manual(
    values = c("cr2" = "goldenrod1",
               "ref_chandler" = "darkseagreen4",
               "term_chandler" = "darkseagreen3"),
    labels = c("cr2" = "E HiFi",
               "ref_chandler" = "Reference",
               "term_chandler" = "Tree")) +
  scale_color_manual(
    values = c("sv" = "grey30",
               "te_5453" = "#eb7287",
               "te_899" = "#7287eb"),
    labels = c("sv" = "Other SV",
               "te_5453" = "5453 bp TE",
               "te_899" = "899bp TE")) +
  scale_y_discrete(
    labels = c("cr2_sv_start" = "E HiFi other SV",
               "cr2_te_5453_start" = "E HiFi 5453 bp",
               "cr2_te_899_start" = "E HiFi 899 bp",
               "ref_chandler_sv_start" = "Reference other SV",
               "term_chandler_sv_start" = "Tree other SV"
    )
  ) +
  labs(x = "Distance from gene start (bp)", fill = "Individual", color = "Type of variant") +
  theme_classic(base_size = 6) +
  theme(
    axis.title.y = element_blank(),
    #axis.text.y = element_blank(),
    #axis.ticks.y = element_blank(),
    legend.key.size = unit(0.2, "cm"),
    legend.position = "none",
    panel.background = element_rect(fill = NA, color = NA),
    plot.background = element_rect(fill = NA, color = NA),
    plot.margin = margin(t = 0.5, r = 0.05, b = 0, l = 0, unit = "cm")
  )

ggsave("./Figures/Transposable_Elements/sv_simulation_v_real_to_gene_start.pdf", height = 2.5, width = 2.95)
ggsave("./Figures/Transposable_Elements/sv_simulation_v_real_to_gene_start.png", height = 2.5, width = 2.95)

prox_sim_ins_means %>%
  filter(LOCATION == "start") %>%
  ggplot(aes(x = DISTANCE, y = UNIQUE)) +
  geom_violin(aes(fill = SOURCE), alpha = 0.75) +
  geom_point(data = prox_real_means_ins[LOCATION == "start"], aes(x = DISTANCE, y = UNIQUE, color = TE)) +
  scale_fill_manual(
    values = c("cr2" = "goldenrod1",
               "ref_chandler" = "darkseagreen4",
               "term_chandler" = "darkseagreen3"),
    labels = c("cr2" = "E HiFi",
               "ref_chandler" = "Reference",
               "term_chandler" = "Tree")) +
  scale_color_manual(
    values = c("sv" = "grey30",
               "te_5453" = "#eb7287",
               "te_899" = "#7287eb"),
    labels = c("sv" = "Other SV",
               "te_5453" = "5453 bp TE",
               "te_899" = "899bp TE")) +
  scale_y_discrete(
    labels = c("cr2_sv_start" = "E HiFi other SV",
               "cr2_te_5453_start" = "E HiFi 5453 bp",
               "cr2_te_899_start" = "E HiFi 899 bp",
               "ref_chandler_sv_start" = "Reference other SV",
               "term_chandler_sv_start" = "Tree other SV"
    )
  ) +
  labs(x = "Distance from gene start (bp)", fill = "Individual", color = "Type of variant") +
  theme_classic(base_size = 6) +
  theme(
    axis.title.y = element_blank(),
    #axis.text.y = element_blank(),
    #axis.ticks.y = element_blank(),
    legend.key.size = unit(0.2, "cm"),
    legend.position = "none",
    panel.background = element_rect(fill = NA, color = NA),
    plot.background = element_rect(fill = NA, color = NA),
    plot.margin = margin(t = 0.5, r = 0.05, b = 0, l = 0, unit = "cm")
  )
```

## Plot proximity to stop
```{r, eval=FALSE}
prox_sim_sv_means %>%
  filter(LOCATION == "stop") %>%
  ggplot(aes(x = DISTANCE, y = UNIQUE)) +
  geom_violin(aes(fill = SOURCE), alpha = 0.75) +
  geom_point(data = prox_real_means[LOCATION == "stop"], aes(x = DISTANCE, y = UNIQUE, color = TE)) +
  scale_fill_manual(
    values = c("cr2" = "goldenrod1",
               "ref_chandler" = "darkseagreen4",
               "term_chandler" = "darkseagreen3"),
    labels = c("cr2" = "E HiFi",
               "ref_chandler" = "Reference",
               "term_chandler" = "Tree")) +
  scale_color_manual(
    values = c("sv" = "grey30",
               "te_5453" = "#eb7287",
               "te_899" = "#7287eb"),
    labels = c("sv" = "Other SV",
               "te_5453" = "5453 bp TE",
               "te_899" = "899bp TE")) +
  labs(x = "Distance from gene end (bp)", fill = "Individual", color = "Type of variant") +
  theme_classic(base_size = 6) +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    legend.key.size = unit(0.2, "cm"),
    legend.position = "none",
    panel.background = element_rect(fill = NA, color = NA),
    plot.background = element_rect(fill = NA, color = NA),
    plot.margin = margin(t = 0.5, r = 0.05, b = 0, l = 0, unit = "cm")
  )

ggsave("./Figures/Transposable_Elements/sv_simulation_v_real_to_gene_stop.pdf", height = 2.5, width = 2.3)
ggsave("./Figures/Transposable_Elements/sv_simulation_v_real_to_gene_stop.png", height = 2.5, width = 2.3)

prox_sim_ins_means %>%
  filter(LOCATION == "stop") %>%
  ggplot(aes(x = DISTANCE, y = UNIQUE)) +
  geom_violin(aes(fill = SOURCE), alpha = 0.75) +
  geom_point(data = prox_real_means_ins[LOCATION == "stop"], aes(x = DISTANCE, y = UNIQUE, color = TE)) +
  scale_fill_manual(
    values = c("cr2" = "goldenrod1",
               "ref_chandler" = "darkseagreen4",
               "term_chandler" = "darkseagreen3"),
    labels = c("cr2" = "E HiFi",
               "ref_chandler" = "Reference",
               "term_chandler" = "Tree")) +
  scale_color_manual(
    values = c("sv" = "grey30",
               "te_5453" = "#eb7287",
               "te_899" = "#7287eb"),
    labels = c("sv" = "Other SV",
               "te_5453" = "5453 bp TE",
               "te_899" = "899bp TE")) +
  labs(x = "Distance from gene end (bp)", fill = "Individual", color = "Type of variant") +
  theme_classic(base_size = 6) +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    legend.key.size = unit(0.2, "cm"),
    legend.position = "none",
    panel.background = element_rect(fill = NA, color = NA),
    plot.background = element_rect(fill = NA, color = NA),
    plot.margin = margin(t = 0.5, r = 0.05, b = 0, l = 0, unit = "cm")
  )
```

## Plot the inside of the 
```{r, eval=FALSE}
ggplot(sim_sv_means, aes(x = GENIC, y = UNIQUE)) +
  geom_violin(aes(fill = SOURCE), alpha = 0.6) +
  geom_point(data = real_means, aes(x = GENIC, y = UNIQUE, color = TE)) +
  scale_fill_manual(
    values = c("cr2" = "goldenrod1",
               "ref_chandler" = "darkseagreen4",
               "term_chandler" = "darkseagreen3"),
    labels = c("cr2" = "E HiFi",
               "ref_chandler" = "Reference",
               "term_chandler" = "Tree")) +
  scale_color_manual(
    values = c("sv" = "grey33",
               "te_899" = "#7287eb",
               "te_5453" = "#eb7287"
    ),
    labels = c("sv" = "Other SV",
               "te_5453" = "5453 bp TE",
               "te_899" = "899 bp TE")) +
  labs(x = "Proportion within gene body", fill = "Simulated distribution", color = "Observed") +
  guides(fill = guide_legend(order = 1), color = guide_legend(order = 2)) +
  theme_classic(base_size = 6) +
  theme(    
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    legend.position = c(0.81,0.91),
    legend.box = "veritcal",
    legend.spacing.y = unit(-0.1, "cm"),
    legend.key.size = unit(0.2, "cm"),
    legend.background = element_rect(fill = NA, color = NA),
    legend.key = element_rect(fill = NA, color = NA),
    panel.background = element_rect(fill = NA, color = NA),
    plot.background = element_rect(fill = NA, color = NA),
    plot.margin = margin(t = 0.5, r = 0.05, b = 0, l = 0, unit = "cm")
  )

ggsave("./Figures/Transposable_Elements/sv_simulation_v_real_in_gene.pdf", height = 2.5, width = 2.3)
ggsave("./Figures/Transposable_Elements/sv_simulation_v_real_in_gene.png", height = 2.5, width = 2.3)

ggplot(sim_ins_means, aes(x = GENIC, y = UNIQUE)) +
  geom_violin(aes(fill = SOURCE), alpha = 0.6) +
  geom_point(data = real_means_ins, aes(x = GENIC, y = UNIQUE, color = TE)) +
  scale_fill_manual(
    values = c("cr2" = "goldenrod1",
               "ref_chandler" = "darkseagreen4",
               "term_chandler" = "darkseagreen3"),
    labels = c("cr2" = "E HiFi",
               "ref_chandler" = "Reference",
               "term_chandler" = "Tree")) +
  scale_color_manual(
    values = c("sv" = "grey33",
               "te_899" = "#7287eb",
               "te_5453" = "#eb7287"
    ),
    labels = c("sv" = "Other SV",
               "te_5453" = "5453 bp TE",
               "te_899" = "899 bp TE")) +
  labs(x = "Proportion within gene body", fill = "Simulated distribution", color = "Observed") +
  guides(fill = guide_legend(order = 1), color = guide_legend(order = 2)) +
  theme_classic(base_size = 6) +
  theme(    
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    legend.position = c(0.81,0.91),
    legend.box = "veritcal",
    legend.spacing.y = unit(-0.1, "cm"),
    legend.key.size = unit(0.2, "cm"),
    legend.background = element_rect(fill = NA, color = NA),
    legend.key = element_rect(fill = NA, color = NA),
    panel.background = element_rect(fill = NA, color = NA),
    plot.background = element_rect(fill = NA, color = NA),
    plot.margin = margin(t = 0.5, r = 0.05, b = 0, l = 0, unit = "cm")
  )
```

# Only plotting areas of interest
## Plot 900 class TEs
```{r, eval=FALSE}
## Start location
prox_sim_ins_means %>%
  filter(LOCATION == "start", TE == "te_899") %>%
  summarise(mean_distance = mean(DISTANCE, na.rm = TRUE))
prox_real_means_ins[LOCATION %in% "start" & TE %in% "te_899"]

prox_sim_ins_means %>%
  filter(LOCATION %in% "start",
         TE %in% "te_899") %>%
  ggplot(aes(x = DISTANCE, y = UNIQUE)) +
  geom_violin(aes(fill = SOURCE), alpha = 0.95, color = "grey50") +
  geom_point(data = prox_real_means_ins[LOCATION %in% "start" & TE %in% "te_899"],
             aes(x = DISTANCE, y = UNIQUE, color = TE),
             size = 2) +
  geom_segment(aes(x = 5974.629, xend = 12316.6, y = 1.12, yend = 1.12), 
               linewidth = 0.25, color = "black") + 
  geom_segment(aes(x = 5974.629, xend = 5974.629, y = 1.1, yend = 1.12), 
               linewidth = 0.25, color = "black") +
  geom_segment(aes(x = 12316.6, xend = 12316.6, y = 1.1, yend = 1.12), 
               linewidth = 0.25, color = "black") + 
  annotate("text", x = (5974.629 + 12316.6) / 2, y = 1.2, 
           label = "0.0051", size = 2)  +
  geom_segment(aes(x = -Inf, xend = Inf, y = Inf, yend = Inf), color = "black", linewidth = 0.5) +
  geom_segment(aes(x = -Inf, xend = -Inf, y = -Inf, yend = Inf), color = "black", linewidth = 0.5) +
  geom_segment(aes(x = -Inf, xend = Inf, y = -Inf, yend = -Inf), color = "black", linewidth = 0.5) +
  scale_fill_manual(
    values = c("cr2" = "grey90"),
    labels = c("cr2" = "E HiFi")) +
  scale_color_manual(
    values = c("te_899" = "#7287eb"),
    labels = c("te_899" = "900 class TE")) +
  labs(x = "Distance from gene start (bp)", fill = "Individual", color = "Type of variant") +
  theme_classic(base_size = 6) +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    legend.key.size = unit(0.2, "cm"),
    legend.position = "none",
    panel.background = element_rect(fill = NA, color = NA),
    plot.background = element_rect(fill = NA, color = NA),
    plot.margin = margin(t = 0.1, r = 0.1, b = 0.1, l = 0.1, unit = "cm")
  )
ggsave("./Figures/Transposable_Elements/ins_simulation_v_real_start_900.pdf", height = 1, width = 1.25)

## stop location
prox_sim_ins_means %>%
  filter(LOCATION == "stop", TE == "te_899") %>%
  summarise(mean_distance = mean(DISTANCE, na.rm = TRUE))
prox_real_means_ins[LOCATION %in% "stop" & TE %in% "te_899"]

prox_sim_ins_means %>%
  filter(LOCATION %in% "stop",
         TE %in% "te_899") %>%
  ggplot(aes(x = DISTANCE, y = UNIQUE)) +
  geom_violin(aes(fill = SOURCE), alpha = 0.95, color = "grey50") +
  geom_point(data = prox_real_means_ins[LOCATION %in% "stop" & TE %in% "te_899"],
             aes(x = DISTANCE, y = UNIQUE, color = TE),
             size = 2) +
  geom_segment(aes(x = 5673.286, xend = 12344.29, y = 1.12, yend = 1.12), 
               linewidth = 0.25, color = "black") + 
  geom_segment(aes(x = 5673.286, xend = 5673.286, y = 1.1, yend = 1.12), 
               linewidth = 0.25, color = "black") + 
  geom_segment(aes(x = 12344.29, xend = 12344.29, y = 1.1, yend = 1.12), 
               linewidth = 0.25, color = "black") +  
  annotate("text", x = (5673.286 + 12344.29) / 2, y = 1.2, 
           label = "0.0031", size = 2)  + 
  geom_segment(aes(x = -Inf, xend = Inf, y = Inf, yend = Inf), color = "black", linewidth = 0.5) +
  geom_segment(aes(x = -Inf, xend = Inf, y = -Inf, yend = -Inf), color = "black", linewidth = 0.5) +
  scale_fill_manual(
    values = c("cr2" = "grey90"),
    labels = c("cr2" = "E HiFi")) +
  scale_color_manual(
    values = c("te_899" = "#7287eb"),
    labels = c("te_899" = "900 class TE")) +
  labs(x = "Distance from gene stop (bp)", fill = "Individual", color = "Type of variant") +
  theme_classic(base_size = 6) +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    legend.key.size = unit(0.2, "cm"),
    legend.position = "none",
    panel.background = element_rect(fill = NA, color = NA),
    plot.background = element_rect(fill = NA, color = NA),
    plot.margin = margin(t = 0.1, r = 0.1, b = 0.1, l = 0.1, unit = "cm")
  )
ggsave("./Figures/Transposable_Elements/ins_simulation_v_real_stop_900.pdf", height = 1, width = 1.25)

## Genic location
sim_ins_means %>%
  filter( TE == "te_899") %>%
  summarise(mean_gene = mean(PROP_SIM_GENIC, na.rm = TRUE))
real_means_ins %>%
  filter( TE == "te_899") %>%
  summarise(mean_gene = mean(PROP_GENIC, na.rm = TRUE))

sim_ins_means %>%
  filter(TE %in% "te_899") %>%
  ggplot(aes(x = GENIC, y = UNIQUE)) +
  geom_violin(aes(fill = SOURCE),
              alpha = 0.95,
              color = "grey50") +
  geom_point(data = real_means_ins[TE %in% "te_899"],
             aes(x = GENIC,
                 y = UNIQUE,
                 color = TE),
             size = 2) +
  geom_segment(aes(x = 0.1714286, xend = 0.2933971, y = 1.12, yend = 1.12), 
               linewidth = 0.25, color = "black") + 
  geom_segment(aes(x = 0.1714286, xend = 0.1714286, y = 1.1, yend = 1.12), 
               linewidth = 0.25, color = "black") + 
  geom_segment(aes(x = 0.2933971, xend = 0.2933971, y = 1.1, yend = 1.12), 
               linewidth = 0.25, color = "black") +  
  annotate("text", x = (0.1714286 + 0.2933971) / 2, y = 1.2, 
           label = "0.0326", size = 2)  + 
  geom_segment(aes(x = -Inf, xend = Inf, y = Inf, yend = Inf), color = "black", linewidth = 0.5) +
  geom_segment(aes(x = Inf, xend = Inf, y = -Inf, yend = Inf), color = "black", linewidth = 0.5) +
  geom_segment(aes(x = -Inf, xend = Inf, y = -Inf, yend = -Inf), color = "black", linewidth = 0.5) +
  scale_fill_manual(
    values = c("cr2" = "grey90"
    ),
    labels = c("cr2" = "Simulation"
    )) +
  scale_color_manual(
    values = c(
      "te_899" = "#7287eb"
    ),
    labels = c(
      "te_899" = "900 class")) +
  scale_x_continuous(breaks = c(0.2, 0.4, 0.6)) +
  labs(x = "Proportion within gene body", fill = "Simulated distribution", color = "Observed") +
  guides(fill = guide_legend(order = 1), color = guide_legend(order = 2)) +
  theme_classic(base_size = 6) +
  theme(    
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    legend.title = element_blank(),
    legend.position = c(0.77,0.82),
    legend.box = "veritcal",
    legend.box.spacing = unit(0, "cm"),
    legend.spacing.y = unit(-0.2, "cm"),
    legend.key.size = unit(0.2, "cm"),
    legend.background = element_rect(fill = NA, color = NA),
    legend.key = element_rect(fill = NA, color = NA),
    panel.background = element_rect(fill = NA, color = NA),
    plot.background = element_rect(fill = NA, color = NA),
    plot.margin = margin(t = 0.1, r = 0.1, b = 0.1, l = 0.1, unit = "cm")
  )
ggsave("./Figures/Transposable_Elements/ins_simulation_v_real_genes_900.pdf", height = 1, width = 1.25)
```

## Plot 5500 class TEs
```{r, eval=FALSE}
## Start location
prox_sim_ins_means %>%
  filter(LOCATION == "start", TE == "te_5453") %>%
  summarise(mean_distance = mean(DISTANCE, na.rm = TRUE))
prox_real_means_ins[LOCATION %in% "start" & TE %in% "te_5453"]

prox_sim_ins_means %>%
  filter(LOCATION %in% "start",
         TE %in% "te_5453") %>%
  ggplot(aes(x = DISTANCE, y = UNIQUE)) +
  geom_violin(aes(fill = SOURCE), alpha = 0.95, color = "grey50") +
  geom_point(data = prox_real_means_ins[LOCATION %in% "start" & TE %in% "te_5453"],
             aes(x = DISTANCE, y = UNIQUE, color = TE),
             size = 2) +
  geom_segment(aes(x = 4389.588, xend = 12740.23, y = 1.12, yend = 1.12), 
               linewidth = 0.25, color = "black") + 
  geom_segment(aes(x = 4389.588, xend = 4389.588, y = 1.1, yend = 1.12), 
               linewidth = 0.25, color = "black") +
  geom_segment(aes(x = 12740.23, xend = 12740.23, y = 1.1, yend = 1.12), 
               linewidth = 0.25, color = "black") + 
  annotate("text", x = (4389.588 + 12740.23) / 2, y = 1.2, 
           label = "0", size = 2)  +
  geom_segment(aes(x = -Inf, xend = Inf, y = Inf, yend = Inf), color = "black", linewidth = 0.5) +
  geom_segment(aes(x = -Inf, xend = -Inf, y = -Inf, yend = Inf), color = "black", linewidth = 0.5) +
  geom_segment(aes(x = -Inf, xend = Inf, y = -Inf, yend = -Inf), color = "black", linewidth = 0.5) +
  scale_fill_manual(
    values = c("cr2" = "grey90"),
    labels = c("cr2" = "E HiFi")) +
  scale_color_manual(
    values = c("te_5453" = "#eb7287"),
    labels = c("te_5453" = "5500 class TE")) +
  labs(x = "Distance from gene start (bp)", fill = "Individual", color = "Type of variant") +
  theme_classic(base_size = 6) +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    legend.key.size = unit(0.2, "cm"),
    legend.position = "none",
    panel.background = element_rect(fill = NA, color = NA),
    plot.background = element_rect(fill = NA, color = NA),
    plot.margin = margin(t = 0.1, r = 0.1, b = 0.1, l = 0.1, unit = "cm")
  )
ggsave("./Figures/Transposable_Elements/ins_simulation_v_real_start_5500.pdf", height = 1, width = 1.25)

## stop location
prox_sim_ins_means %>%
  filter(LOCATION == "stop", TE == "te_5453") %>%
  summarise(mean_distance = mean(DISTANCE, na.rm = TRUE))
prox_real_means_ins[LOCATION %in% "stop" & TE %in% "te_5453"]

prox_sim_ins_means %>%
  filter(LOCATION %in% "stop",
         TE %in% "te_5453") %>%
  ggplot(aes(x = DISTANCE, y = UNIQUE)) +
  geom_violin(aes(fill = SOURCE), alpha = 0.95, color = "grey50") +
  geom_point(data = prox_real_means_ins[LOCATION %in% "stop" & TE %in% "te_5453"],
             aes(x = DISTANCE, y = UNIQUE, color = TE),
             size = 2) +
  geom_segment(aes(x = 4614.191, xend = 12722.25, y = 1.12, yend = 1.12), 
               linewidth = 0.25, color = "black") + 
  geom_segment(aes(x = 4614.191, xend = 4614.191, y = 1.1, yend = 1.12), 
               linewidth = 0.25, color = "black") + 
  geom_segment(aes(x = 12722.25, xend = 12722.25, y = 1.1, yend = 1.12), 
               linewidth = 0.25, color = "black") +  
  annotate("text", x = (4614.191 + 12722.25) / 2, y = 1.2, 
           label = "0", size = 2)  + 
  geom_segment(aes(x = -Inf, xend = Inf, y = Inf, yend = Inf), color = "black", linewidth = 0.5) +
  geom_segment(aes(x = -Inf, xend = Inf, y = -Inf, yend = -Inf), color = "black", linewidth = 0.5) +
  scale_fill_manual(
    values = c("cr2" = "grey90"),
    labels = c("cr2" = "E HiFi")) +
  scale_color_manual(
    values = c("te_5453" = "#eb7287"),
    labels = c("te_5453" = "5500 class TE")) +
  labs(x = "Distance from gene stop (bp)", fill = "Individual", color = "Type of variant") +
  theme_classic(base_size = 6) +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    legend.key.size = unit(0.2, "cm"),
    legend.position = "none",
    panel.background = element_rect(fill = NA, color = NA),
    plot.background = element_rect(fill = NA, color = NA),
    plot.margin = margin(t = 0.1, r = 0.1, b = 0.1, l = 0.1, unit = "cm")
  )
ggsave("./Figures/Transposable_Elements/ins_simulation_v_real_stop_5500.pdf", height = 1, width = 1.25)

## Genic location
sim_ins_means %>%
  filter( TE == "te_5453") %>%
  summarise(mean_gene = mean(PROP_SIM_GENIC, na.rm = TRUE))
real_means_ins %>%
  filter( TE == "te_5453") %>%
  summarise(mean_gene = mean(PROP_GENIC, na.rm = TRUE))

sim_ins_means %>%
  filter(TE %in% "te_5453") %>%
  ggplot(aes(x = GENIC, y = UNIQUE)) +
  geom_violin(aes(fill = SOURCE),
              alpha = 0.95,
              color = "grey50") +
  geom_point(data = real_means_ins[TE %in% "te_5453"],
             aes(x = GENIC,
                 y = UNIQUE,
                 color = TE)) +
  geom_segment(aes(x = 0.519084, xend = 0.2803687, y = 1.12, yend = 1.12), 
               linewidth = 0.25, color = "black") + 
  geom_segment(aes(x = 0.519084, xend = 0.519084, y = 1.1, yend = 1.12), 
               linewidth = 0.25, color = "black") + 
  geom_segment(aes(x = 0.2803687, xend = 0.2803687, y = 1.1, yend = 1.12), 
               linewidth = 0.25, color = "black") +  
  annotate("text", x = (0.519084 + 0.2803687) / 2, y = 1.2, 
           label = "0", size = 2)  + 
  geom_segment(aes(x = -Inf, xend = Inf, y = Inf, yend = Inf), color = "black", linewidth = 0.5) +
  geom_segment(aes(x = Inf, xend = Inf, y = -Inf, yend = Inf), color = "black", linewidth = 0.5) +
  geom_segment(aes(x = -Inf, xend = Inf, y = -Inf, yend = -Inf), color = "black", linewidth = 0.5) +
  scale_fill_manual(
    values = c("cr2" = "grey90"
    ),
    labels = c("cr2" = "Simulation"
    )) +
  scale_color_manual(
    values = c(
      "te_5453" = "#eb7287"
    ),
    labels = c(
      "te_5453" = "5500 class")) +
  scale_x_continuous(breaks = c(0.2, 0.4, 0.6)) +
  labs(x = "Proportion within gene body", fill = "Simulated distribution", color = "Observed") +
  guides(fill = guide_legend(order = 1), color = guide_legend(order = 2)) +
  theme_classic(base_size = 6) +
  theme(    
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    legend.title = element_blank(),
    legend.position = c(0.77,0.82),
    legend.box = "veritcal",
    legend.box.spacing = unit(0, "cm"),
    legend.spacing.y = unit(-0.2, "cm"),
    legend.key.size = unit(0.2, "cm"),
    legend.background = element_rect(fill = NA, color = NA),
    legend.key = element_rect(fill = NA, color = NA),
    panel.background = element_rect(fill = NA, color = NA),
    plot.background = element_rect(fill = NA, color = NA),
    plot.margin = margin(t = 0.1, r = 0.1, b = 0.1, l = 0.1, unit = "cm")
  )
ggsave("./Figures/Transposable_Elements/ins_simulation_v_real_genes_5500.pdf", height = 1, width = 1.25)
```